---
title: "Data processing for chapter 3: 10-45 m sites" MKMRS data of fish survey only
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
 A. Fish_abund df 
 C. factors3 df
 D. Fish_Inv_fact combined by "tag" df 
---

### step 1 load data library etc and set wd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
library(assertthat)
library(devtools)
library(ggridges)

library(ggtext)# for italics in the plot legend 
library(here)

# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")

# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
complex.inic <- read.csv(here("R.generated.Data", "HOBO_compl_45.csv"))


fish_data <-merge(fish_data,sites, by= "site_id") 

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data$species[fish_data$species == 'Diplodos cervinus'] <- 'Diplodus cervinus'
fish_data$species[fish_data$species == 'Salpa salpa'] <- 'Sarpa salpa'

fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))

head(fish_data)
```

```{r}
#change date to char,and summarize
fish_date_site_species <-fish_data %>%
  mutate(date_char = as.character(sample_date))

head (fish_date_site_species)

```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs

```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter<-fish_date_site_species %>% filter((site=="Achziv" | site=="SdotYam"))
 
fish_filter <-   fish_filter%>%
  filter(distance < 5.1)



head(fish_filter)
Visibility <- fish_filter %>%
  select(12,22) %>%
 unique()
head(Visibility)
```

###making 3 column abudance data

```{r}

com_mat<-fish_filter%>%
   dplyr::select(15,22)
com_mat<-mutate(com_mat, occ =1 ) 

occ_int<-as.integer(com_mat$occ)

head(com_mat)
```

###making Fish abudance matrix

```{r}

SumBySpecies<- ddply(com_mat, c("tag","species"), summarise,
               occurance = sum(occ))
head(SumBySpecies)

#There are 2 ways. either use 
SumBySpecies_mat <-matrify(SumBySpecies)
distinct(SumBySpecies_mat)
#or the next:
#SumBySpecies_mat <-  dcast(SumBySpecies,tag~species)
head(SumBySpecies_mat)


```

###making Fish abudance table filtering empty species

```{r}
Fish_abund<-as.data.frame(SumBySpecies_mat)

Fish_abund_filtered <- Fish_abund[, colSums(Fish_abund != 0) > 0]




```
### Fish Abundance matrix : all_data_clean 
```{r}
#Change NAs to zero
 
all_data_clean <- Fish_abund_filtered
all_data_clean[is.na(all_data_clean)] = 0 


head(all_data_clean)

not_empty(all_data_clean[0, ])
not_empty(all_data_clean[, 0])

write.csv(all_data_clean, here("R.generated.Data", "45Fish_data_clean.csv"))


```
### Invertebrate summary according to taxonomy_group : allFish10.45dataFact


```{r}

Inv.Grp.data <- read.csv(here( "Data","inv_group_mat_wis.csv"))
Inv.Grp.data = subset(Inv.Grp.data, select = -c(tag) )
Inv.Grp.data<- Inv.Grp.data %>% mutate (tag= paste0(site
,depth,transect,season,year))
head(Inv.Grp.data)
# Add a new column with the row names
allFish10.45data <-all_data_clean

library(tibble)

# Convert row names to a column
Fish_abund_filtered <- rownames_to_column(Fish_abund_filtered, var = "RowID")
colnames(Fish_abund_filtered)[colnames(Fish_abund_filtered) == "RowID"] <- "tag"

colnames(Inv.Grp.data)
Inv.Grp.factors <-select(Inv.Grp.data,Algae,Bryozoan,Hard.Substrate, Hydrozoa, Soft.Substrate,Sponge,tag, 11:19)

head(Inv.Grp.factors)

Inv.Grp.factors <-as.data.frame(Inv.Grp.factors)
Fish_abund_filtered <-as.data.frame(Fish_abund_filtered)

head(allFish10.45dataFact)
DepthFishdataFact <-select(allFish10.45dataFact, 1, 56:69)
 #there are errors in the "Reserve" column

write.csv(Fish_abund_filtered, here("R.generated.Data","allFish10.45data.csv"))

anyDuplicated(names(Fish_abund_filtered))
anyDuplicated(names(Inv.Grp.factors))

```
###Combine visibility and complexity factors into the DepthFishdataFact df
```{r}

DepthFishdataFact <-left_join(DepthFishdataFact, Visibility, by = "tag")
DepthFishdataFact <-left_join(DepthFishdataFact, complex.inic, by = "tag")
DepthFishdataFact = subset(DepthFishdataFact, select = -c(X) )
write.csv(DepthFishdataFact, here("R.generated.Data","Fish.FactorsGrp.wis.csv"))
```

