---
title: "Diversity"
author: "Belmaker Lab"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(tidyverse)
library(plotrix)
library(betapart)
library(here)
```
```{r}
my_data <- read.csv(here("R.generated.Data","Fish.FactorsGrp.wis.csv")) #change to your file directory
 head(my_data)                  

my_data<- subset(my_data, select = -c(X) )
colnames(my_data)
my_data$site_id <- factor(my_data$site_id,ordered = T)
my_data<-my_data%>%
  mutate(Year_season_id =paste0(sample_year,season,site_id)) %>%
  relocate(74,.after = tag) 
head(my_data)

species_dist<- my_data %>% select ( "Epinephelus.marginatus" , "Mycteroperca.rubra" ,"Siganus.luridus" ,   "Epinephelus.costae" ,  "Siganus.rivulatus")

# Assuming your dataset is named "my_data" and you want to summarize columns "visibility", "temperature", and "humidity"

# List of columns to summarize
columns_to_summarize <- c("Epinephelus.sp","Epinephelus.marginatus","Mycteroperca.rubra","Siganus.luridus","Epinephelus.costae","Siganus.rivulatus","Siganus.sp")

# Initialize an empty list to store distributions
distributions <- list()

# Calculate distribution for each column
for (col in columns_to_summarize) {
  distributions[[col]] <- table(my_data[[col]])
}

# Print distributions
for (col in columns_to_summarize) {
  cat("Distribution for column", col, ":\n")
  print(distributions[[col]])
}

```
```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpT <- ggplot(my_data, aes(x = site_id, y = LogVR)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Log(Vertical Relief)")

# Set y-axis limits without truncating data points
bxpT <- bxpT + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpT)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpT <- bxpT     + labs(x = NULL)   # 
bxpT
```

```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpS <- ggplot(my_data, aes(x = site_id, y = Rugosity)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Rugosity")

# Set y-axis limits without truncating data points
#bxpS <- bxpS + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpS)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpS <- bxpS     + labs(x = NULL)   # 
bxpS
```


```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpH <- ggplot(my_data, aes(x = site_id, y = Slope)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Slope")

# Set y-axis limits without truncating data points
#bxpH <- bxpH + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpH)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpH <- bxpH     + labs(x = NULL)   # 
bxpH
```
```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpC <- ggplot(my_data, aes(x = site_id, y = Curvature)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Curvature")

# Set y-axis limits without truncating data points
#bxpC <- bxpC + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpC)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpC <- bxpC     + labs(x = NULL)   # 
bxpC
```
### ANOVA testing
```{r}
# Convert depth_category to factor for correct analysis
my_data$depth_category <- as.factor(my_data$depth_category)
my_data_sy <-my_data %>%
  filter(site == "SdotYam")
# Perform two-way ANOVA
anova_result <- aov(LogVR ~ site + depth_category , data = my_data)

# Summarize the ANOVA results
summary(anova_result)

# Perform two-way ANOVA
anova_result <- aov(LogVR ~ site + depth_category , data = my_data)

# Summarize the ANOVA results
summary(anova_result)

# Perform one-way ANOVA for SY
anova_result <- aov(LogVR ~ site + depth_category , data = my_data)

```


```{r}
library(ggpubr)
# Set a common x-axis titlec ommon.legend = TRUE,  common.xlabel = "N" )
tiff("Habitat heterogeneity indices.jpeg", units="in", width=7, height=6, res=600)
diversPlots <- ggarrange( bxpT, bxpS, bxpH, bxpC, 
          labels = c("       a ", "      b ", "       c ", "      d   "),
          ncol = 2, nrow = 2)   # Set a common x-axis title)
diversPlots
dev.off()
diversPlots

```
###Species Matrix +Metadata
```{r}
first_species <- 30
sp_matrix <- my_data[,first_species:ncol(my_data)]
meta_data <- my_data[,1:30] # columns i want to keep from my data



```
### Hill numbers profile Allways use  hill = T
```{r}
renyi_profile <- renyi(sp_matrix,  
                       scales = c(0, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 4, 8, 16, 32, 64, Inf),
                       hill = T)
head(renyi_profile)

renyi_df<-bind_cols(meta_data,renyi_profile)

renyi_df<-gather(renyi_df,"Q","Value",30:44)
head(renyi_df)
renyi_df<-renyi_df %>% arrange(site_id,depth_category,sample_year, season,tag) %>%
   filter(!is.na(Q))

renyi_df$Q <- as.numeric(renyi_df$Q)
```

```{r}
head(renyi_df)
ggplot( data = renyi_df ) +
  aes(x = Q, y = Value, group = tag , color = -depth_category)+
  geom_line() +
   scale_x_continuous(limits = c(0,10)) + scale_y_continuous(limits = c(0,13)) +facet_wrap(.~site_id) +
         labs(y = expression("Diversity  D(p) "^"Q")) +theme_classic()

ggplot(data = renyi_df) +
  aes(x = Q, y = Value, group = tag, color = factor(site_id)) +
  geom_line() +
  scale_x_continuous(limits = c(0, 10)) +
  scale_y_continuous(limits = c(0, 13)) +
  facet_wrap(. ~ site_id) +
  labs(y = expression("Diversity  D(p) "^"Q")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) +
  scale_color_manual(values = c("lightblue", "blue", "darkorange", "brown"))  # Specify color palette



```


###SEction 2: Grouping per sampling day
```{r}
colnames(my_data)
group_year_data <- my_data %>% 
  dplyr::select(sample_date, Year_season_id, site_id, depth_category, season, 30:74) %>% 
  dplyr::group_by(sample_date, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across(Acantholabrus.palloni:Torquigener.flavimaculosus, sum),.groups = "keep")
head(group_year_data)



```


```{r}

renyi_profile_group <- renyi(group_year_data[,-c(1:5)],
                             scales = c(0, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 4, 8, 16, 32, 64, Inf),  hill = T)

renyi_df_group<-bind_cols(group_year_data[,1:5],renyi_profile_group)

#compose long df with  new Q grouping
renyi_df_group<-gather(renyi_df_group,"Q","Value",6:20)

renyi_df_group<-renyi_df_group %>% arrange(Year_season_id, site_id,depth_category, season )

renyi_df_group$Q <- as.numeric(renyi_df_group$Q)
renyi_df_group$depth_category <- as.factor(renyi_df_group$depth_category)
head(renyi_df_group)
```

```{r}
ggplot(renyi_df_group,aes(x=Q, y=Value, group = Year_season_id, color= site_id))+ 
  geom_line()+
  scale_x_continuous(limits = c(0,15))+facet_grid (~    site_id +depth_category)+ theme_classic() +
         labs(y = expression("Diversity  D(p) "^"Q"))+
  scale_color_manual(values = c("lightblue", "blue", "darkorange", "brown"))  # Specify color palette
```
#clean my environment so it will be less clutter and messy… this is the line that remove all object (except my_data and first_species) from my environment:
```{r}
#rm(list=setdiff(ls(), c("my_data", "group_year_data","first_species")))
```

###Pairwise  Bray-Curtis similarity

# ```{r}
# 
# Fall_data <- my_data %>% filter(season == "Spring", site_id == "SdotYam45") # filter for example
# sp_matrix <- Fall_data[,first_species:ncol(Fall_data)]
# bray_results <- bray.part(sp_matrix)
# mean(bray_results[[3]])
# std.error(bray_results[[3]])
# ```
###Loop of pairwise comparisons 
```{r}


bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(group_year_data$site_id)){
  
site_data <- group_year_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +ylab(" Bray-Curtis dissimilarity")+
  scale_color_manual(values = c("lightblue", "blue", "darkorange", "brown"))  # Specify color palette


```
#One site, SdotYam25 Fall and plot it vs time

```{r}

one_site<-my_data %>% filter(site_id == 'Achziv45', season== "Fall") %>%
  arrange(sample_date)

# Convert sample_date to Date object
one_site$sample_date <- as.Date(one_site$sample_date, format = "%d/%m/%Y")

first_species <-31

bray_scores_SY25<-list() # create an empty list to store my newly created data

for (i in unique(one_site$sample_date)){
  
knoll_data <- one_site %>% filter(sample_date == i) # keep only the observation of sample i

knoll_sp_matrix <- knoll_data[,first_species:length(knoll_data)] # create species matrix

knoll_bray_part <- bray.part(knoll_sp_matrix) # apply the function that calculate bray Curtis distances 
knoll_bray_results<- knoll_bray_part[[3]] # keep only the bray Curtis results

knoll_bray_results<-as.numeric(knoll_bray_results) # convert to numeric object

mean_bray <- mean(knoll_bray_results)
se_bray <- std.error(knoll_bray_results)

sample_date <- i

bray_data_knoll <- data.frame(sample_date,mean_bray,se_bray)

bray_scores_SY25[[i]]<-bray_data_knoll
  
}

bray_scores_SY25<-bind_rows(bray_scores_SY25)
#change from serial to date
bray_scores_SY25$sample_date<-as.Date(bray_scores_SY25$sample_date, origin = "1970-01-01")


SdotYam25FallP <- ggplot(bray_scores_SY25,aes(x = sample_date,
                       y = mean_bray,
                       color = sample_date)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2)

SdotYam25FallP + 
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 year") +
  ylab(" Bray-Curtis dissimilarity") +
  xlab("Survey Years") +
  ggtitle( "Achziv 45m depth site community dissimilarity with time" ) +theme_classic() +
  theme(legend.position = "none")


```
###Log transformed group data


###Species Matrix +Metadata
```{r}
head(my_data)
first_species <- 28



```

```{r}

my_data_log<-decostand(sp_matrix, method="hellinger")
```
###Bray scores per site: transformed data hellinger
```{r}
bray_scores_log_10<-list() # create an empty list to store my newly created data

for (i in unique(my_data_log$site_id)){
  
year_data <- my_data_log %>% filter(site_id == i) # keep only the observation of sample i

year_sp_matrix <- year_data[,first_species:length(year_data)] # create species matrix

year_sp_matrix_log<-decostand(year_sp_matrix,method = 'hellinger')

year_bray_part_log <- bray.part(year_sp_matrix_log) # apply the function that calculate bray Curtis distances 

year_bray_results_log<- year_bray_part_log[[3]] # keep only the bray Curtis results

year_bray_results_log<-as.numeric(year_bray_results_log) # convert to numeric object

mean_bray <- mean(year_bray_results_log) # calculate the mean bray curtis distance
se_bray <- std.error(year_bray_results_log)# calculate SE
Sample <- i # argument with the the name of the site

bray_data_log <- data.frame(Sample,mean_bray,se_bray) # create data frame that save those variables 

bray_scores_log_10[[i]]<-bray_data_log # save it in my list
  
}

bray_scores_hell<-bind_rows(bray_scores_log_10) # convert from list to data frame

# get the Years order right

#bray_scores_log_10$Sample <- factor(bray_scores_log_10$Sample,levels = c("2018 Fall",                                         "2020 Spring",                                         "2020 Fall",                                         "2021 Spring"),ordered = T)

# lets plot:

ggplot(bray_scores_hell,aes(x = Sample,
                       y = mean_bray,
                       color = Sample)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2)+
  ggtitle("Log10 transformed data")
```

###  BETA diversity components:  depth example


```{r}
one_site<-my_data %>% #filter(site_id ==  "Achziv45" ) %>%
  arrange(sample_date)
  #'Achziv10'|site_id ==  "Achziv25" | site_id
  #filter(season == "Fall") %>%
  

# Convert sample_date to Date object
one_site$sample_date <- as.Date(one_site$sample_date, format = "%d/%m/%Y")
head(one_site)
#Ach sites sp. matrix

sp_matrix <- one_site[,first_species:ncol(one_site)]
meta_data <- one_site[,1:6] # columns i want to keep from my data

```
##Incidence data- beta pairs
```{r}
AC_incidence <- sp_matrix %>% 
  replace(sp_matrix > 0, 1) 
AC_incidence_bind<-bind_cols(meta_data,AC_incidence )


caves_beta_pairs <- beta.pair(AC_incidence)
```
We get 3 distance matrixes in one list:

The distance which is due to species turnover between sites.
The distance which is due to species nestedness between sites.
The total distance (nestedness + turnonver).

IMPORTANT NOTE: unlike the pairwise comparisons multi-site comparisons are sensitive to sample size! in order for the analyses to work properly you need to have equal sample in each group you want to compare! If you don’t have equal sample size in each category, continue to the next section!


```{r}
summary_table<-one_site %>% group_by(sample_year, season, depth_category, site_id) %>% summarise("tags"= n_distinct(tag))
summary_table
```
Sample size is Good!


```{r}
same_effort_beta<-list()


for (i in AC_incidence_bind$depth_category) {
  
  one_year <- AC_incidence_bind %>% filter(depth_category == i)
  
  one_year_sp_matrix <- one_year[,first_species:ncol(one_year)]
  
  beta_multi <- as.data.frame(beta.multi(one_year_sp_matrix))
  
  beta_multi$depth_category<-i
  
  same_effort_beta[[i]]<-beta_multi
  
  
}



same_effort_beta<-bind_rows(same_effort_beta)
#same_effort_beta$site_id <- as.Date(same_effort_beta$site_id,   origin = "1970-01-01")

same_effort_beta<- gather(same_effort_beta,"component","beta",1:3)

same_effort_beta <- same_effort_beta %>%
  filter(component != "beta.SOR") 

ggplot(same_effort_beta)+
  aes(x=depth_category,y=beta,fill=component)+
  geom_bar(stat="identity",position = position_stack())+
  ylab(expression(paste(beta,"-diversity"))) +
  #scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "2 year") +
  ylab(" Bray-Curtis dissimilarity") +
  xlab("Survey sites") +
  ggtitle( "Survey sites pooled 2015-2022 community dissimilarity" ) +theme_classic() 
```

```{r}
library(iNEXT)
data(spider)
head(spider)
spideriN <-iNEXT(spider, q=c(0,1,2), datatype="abundance")
speciesiN <- iNEXT(sp_matrix,  q=c(0,1,2), datatype="abundance")
ggiNEXT(sp_matrix, type=3, se=TRUE, grey=FALSE,)
```

