---
title: "Data processing for chapter 3: 25-45 m sites" MKMRS data of fish survey  and invertebrate photoquadrates
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
WHEN UPDATING A NEW FISH_SURVEY FILE- CORRECT VIVIBILITY TO 7 AT AchzivR453Spring2020
AchzivR454Spring2020

---

### step 1 load data library etc and set wd


```{r setup, include=FALSE}



library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
library(assertthat)
library(devtools)
#For Diversity calculations
library(plotrix)
library(betapart)
library(here)
library(ggridges)

library(ggtext)# for italics in the plot legend 
library(ggpubr) 
# Richness
library(rareNMtests)
library(mobr)
here<-here::here


```

```{r}
# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")
csv_path2 <- here("data", "45FBsize_a_b_with_missing.csv")
# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
FBsize <- read.csv(csv_path2)
complex.inic <-read.csv(here("data", "EnvVar10_45.csv"))
Diet <- read_csv(here("data","Diet.csv"))
fish_data <-merge(fish_data,sites, by= "site_id") 
trait <-read.csv(here("data", "Belmaker_Species_Traits.csv"))

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))
#Make sure Achziv is Reserve and Sdot Yam Not
head(trait)
MedSp <-select(trait,"Species_updated_name_fishbase", "Lessepsian")

head(fish_data)
colnames(MedSp)<-c("species", "Lessepsian")
MedSp$Lessepsian <-   as.factor(MedSp$Lessepsian)


```


###QA
```{r}

# Group by "tag" and count unique "sample_date" values
result <- fish_data %>%
  group_by(tag) %>%
  summarise(unique_dates = n_distinct(sample_date)) %>%
  filter(unique_dates > 1)

# Print the result
print(result)
```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs
Filtering species with Fishbase biomass constants
###Renaming fish name spelling errors; renaming reserve status; cropping distance from transect 
```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter <- fish_data %>%
  filter(
    depth_category > 10 & depth_category < 40 &
    (sample_year > 2019) & (sample_year < 2022) &
    !site_id %in% c("Nahariya", "Ashdod") & season=="Fall" 
  )

# Get the unique species names from the FBsize data frame
FB_species <- unique(FBsize$Fbspecies_chosen)
fish_filter$species <- sub("Salpa salpa", "Sarpa salpa", fish_filter$species)
fish_filter$species <- sub("Saprisoma cretense", "Sparisoma cretense", fish_filter$species)
fish_filter$species <- sub("Diplodos cervinus", "Diplodus cervinus", fish_filter$species)

# Filter the fish_data data frame to keep only rows where the species is present in unique_species

fish_filter <- fish_filter[fish_data$species %in% FB_species, ] %>%
 dplyr:: filter(distance < 3.5)

fish_filter$reserve <- ifelse(fish_filter$site == "Achziv", "R", "N")  
head(fish_filter)

# Filter season
#fish_filter <-   fish_filter%>%
  #filter(season == "Fall")
```

```{r}
fb2use<-FBsize
# Using $ operator with negative indexing
fb2use$Locality <- NULL
fb2use$X <- NULL
fb2use$X.1 <- NULL
head(fb2use)
fb2use <- rename(fb2use, species = Fbspecies_chosen)

fish_filter_a_b <- left_join(fish_filter,fb2use, by= "species")

# Calculate the average size for each species and fill NA values with these averages
fish_filter_a_b <- fish_filter_a_b %>%
  group_by(species) %>%
  mutate(size = ifelse(is.na(size), mean(size, na.rm = TRUE), size)) %>%
  ungroup() %>%
  mutate(fish_weight = a * size^b)

# Display the first few rows
head(fish_filter_a_b)

```
#A. calculate the total number of individuals per species 
norm weight: scale the weight : weight -mean weight devided by the standard deviation

```{r}

fish_long <- fish_filter_a_b %>%
  dplyr::select(-transect, -a_b, -a, -b, -habitat, -depth, -factor_visibility, -observer,  -distance, -visibility, -size, -taxonomy_group) %>%
  arrange(tag)%>% 
  relocate(species, .after = tag) %>%
  dplyr::mutate(
    occ = 1)
#      ) %>% 
#   dplyr::group_by(across(site_id:tag)) %>% 
#   dplyr::mutate(
#        "tag_weight" = sum(fish_weight),  "total_ind" = sum(occ)
#   ) %>% 
#   ungroup()
head(fish_long)

fish_long_med <- left_join(fish_long,MedSp, by = "species")
  # Summarize fish_weight and occ per tag
fish_summary_les <- fish_long_med %>%
   relocate(Lessepsian, .after = tag) %>% 
  group_by((across(site_id:Lessepsian))) %>%
  summarize(
    tag_weight = sum(fish_weight, na.rm = TRUE),  # Summing fish_weight per tag
    total_ind = sum(occ, na.rm = TRUE)            # Summing occ per tag
  )

# View the summary
head(fish_summary_les)
```
#Overall biomass
```{r}
fish_summary <- fish_long  %>% 
  group_by((across(site_id:tag))) %>%
  summarize(
    tag_weight = sum(fish_weight, na.rm = TRUE),  # Summing fish_weight per tag
    total_ind = sum(occ, na.rm = TRUE)            # Summing occ per tag
  )
Biomass <- fish_summary%>%
  arrange(lat)
  
 
Plot_biomass <- Biomass%>%
  ggplot(aes(x = as.factor(lat),y = log2(tag_weight /1000))) +
  geom_boxplot(notch=TRUE, notchwidth = 0.8) +
  #facet_wrap(~lat) +
  theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Log2 of biomass [Kg] per 150 square m") + 
  theme(text = element_text(size = 18)) +
  xlab("Latitude")# +theme(axis.text.x = element_blank())   # Set font size for all text elements
Plot_biomass


```

#Overall abundannce
```{r}
abundannce <- fish_summary 
Plot_abundannce <- abundannce%>%
  ggplot(aes(y = total_ind, x= as.factor(lat))) +
  geom_boxplot(notch=TRUE, notchwidth = 0.8) +
    theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Fish abundance per 150 square m") +  xlab("Latitude") + 
  theme(text = element_text(size = 18))
Plot_abundannce


```
#Lesspesians
```{r}
Plot_biomassLess <- fish_summary_les %>%
  ggplot(aes(y = log2(tag_weight / 1000), x = as.factor(lat), color = Lessepsian)) +  # Use 'color' to partition by Lessepsian
  geom_boxplot() +
 # facet_wrap(~site_id) +
    ylab("Log2 of biomass [Kg] per 150 square m") +
theme_classic() +
  scale_color_manual(values = c("0" = "blue", "1" = "red")) + xlab("Latitude") + 
  theme(text = element_text(size = 18))   # Reverse color scale: 0 = blue, 1 = red

Plot_biomassLess
```

#Lesspesian Abundance
```{r}
Plot_abundLess <- fish_summary_les %>%
  ggplot(aes(y = total_ind, x= as.factor(lat), color = Lessepsian)) +  # Use 'color' to partition by Lessepsian and treat it as a factor
  geom_boxplot(notch = TRUE, notchwidth = 0.3) +
  ylab("Fish abundance per 150 square m") +
  theme_classic() +
  scale_color_manual(values = c("0" = "blue", "1" = "red"))+ 
  theme(text = element_text(size = 18)) + xlab("Latitude")   # Reverse color scale: 0 = blue, 1 = red

Plot_abundLess

```
### Combining with functional grouping
```{r}
fish_long_Diet <- fish_long_med %>% 
  left_join(Diet, by = "species" )  %>%
  filter(!is.na(Diet))
head(fish_long_Diet)
write.csv(fish_long_Diet, here("R.generated.Data","fish_size_diet.csv" ))
```
### Plotting
#B. Calculate  the total number of species (n_species_sample)and total individuals per survey (n_ind_sample))

```{r}
head(fish_summaryDiet)
  fish_summaryDiet <- fish_long_Diet %>%
   relocate(Diet, .after = tag) %>% 
  group_by((across(site_id:Diet))) %>%
  summarize(
    tag_weight = sum(fish_weight, na.rm = TRUE),  # Summing fish_weight per tag
    total_ind = sum(occ, na.rm = TRUE)            # Summing occ per tag
  )

# View the summary
head(fish_summaryDiet)
```
```{r}
# Define the color palette for diets
colorDiet <- c(
  "FC" = "red",
  "H" = "green4",
  "Mi" = "darkorange",
  "Ma" = "brown",
  "PK" = "darkblue"
  )

# Assuming your dataframe is named df
# Ensure that the site_id and Diet columns are factors
fish_summaryDiet <- fish_summaryDiet %>%
  filter(!is.na(Diet)) %>%
  mutate(
    site_id = factor(site_id),
    Diet = factor(Diet, levels = names(colorDiet))  # This ensures the order of diet categories
  )

# Create the stacked bar plot
ggplot(data = fish_summaryDiet, aes(x = as.factor(lat), y = total_ind, fill = Diet)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = colorDiet) +  # Apply the custom color palette
  theme_classic() +
  labs(
    x = "Site ID",
    y = "Fish Abundance",
    fill = "Diet"
  ) +
  theme(
    text = element_text(size = 18),  # Set overall text size to 18
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
  )
```

##plotting the relationship between species and individuals number:
```{r}
colorDiet = c("red", "green4", "darkorange","brown","darkblue")
fish_summaryDiet  %>%
  ggplot(aes(x=log10(total_ind),
           y=log2(tag_weight/1000),
           color = Diet, shape = Lessepsian))+
        xlab("Log10(Trophic abundance per sampling day)")+
       ylab("Log2(Biomass, Kg)")+
      ggtitle("Relationship between Trophic group abundance and biomass")+
  geom_jitter()+
    facet_wrap(~site_id ) +theme_classic()  +
  scale_color_manual(values = colorDiet) # Specify color palette
```
```{r}
colorDiet = c("red", "green4", "darkorange","brown","grey", "darkblue")
fish_summaryDiet %>%
  filter(site == "Achziv" | site == "SdotYam") %>%
  ggplot(aes(x=log10(total_ind),
           y=log2(tag_weight/1000),
           color = Diet))+
        xlab("Log10(Trophic abundance per sampling day)")+
       ylab("Log2(Biomass, Kg)")+
      ggtitle("Relationship between Trophic group abundance and biomass")+
  geom_jitter()+
    facet_wrap(~site_id ) +theme_classic()  +
  scale_color_manual(values = colorDiet) # Specify color palette
```