---
title: "Data processing for chapter 3: 25-45 m sites" MKMRS data of fish survey  and invertebrate photoquadrates
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
WHEN UPDATING A NEW FISH_SURVEY FILE- CORRECT VIVIBILITY TO 7 AT AchzivR453Spring2020
AchzivR454Spring2020

---

### step 1 load data library etc and set wd


```{r setup, include=FALSE}



library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
library(assertthat)
library(devtools)
#For Diversity calculations
library(plotrix)
library(betapart)
library(here)
library(ggridges)

library(ggtext)# for italics in the plot legend 
library(ggpubr) 
# Richness
library(rareNMtests)
library(mobr)
here<-here::here


```

```{r}
# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")
csv_path2 <- here("data", "45FBsize_a_b_with_missing.csv")
# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
FBsize <- read.csv(csv_path2)
complex.inic <-read.csv(here("data", "EnvVar10_45.csv"))
Diet <- read_csv(here("data","Diet.csv"))
trait <-read.csv(here("data", "Belmaker_Species_Traits.csv"))

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))
#Make sure Achziv is Reserve and Sdot Yam Not
head(trait)
MedSp <-select(trait,"Species_updated_name_fishbase", "Lessepsian")

head(fish_data)
colnames(MedSp)<-c("species", "Lessepsian")
MedSp$Lessepsian <-   as.factor(MedSp$Lessepsian)
color6 = c("lightblue3", "blue2", "darkblue","yellow3", "darkorange", "brown")

```


###QA
```{r}

# Group by "tag" and count unique "sample_date" values
result <- fish_data %>%
  group_by(tag) %>%
  summarise(unique_dates = n_distinct(sample_date)) %>%
  filter(unique_dates > 1)

# Print the result
print(result)
```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs
Filtering species with Fishbase biomass constants
###Renaming fish name spelling errors; renaming reserve status; cropping distance from transect 
```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter<-fish_data %>%dplyr:: filter((site=="Achziv" | site=="SdotYam"))
# Get the unique species names from the FBsize data frame
FB_species <- unique(FBsize$Fbspecies_chosen)
fish_filter$species <- sub("Salpa salpa", "Sarpa salpa", fish_filter$species)
fish_filter$species <- sub("Saprisoma cretense", "Sparisoma cretense", fish_filter$species)
fish_filter$species <- sub("Diplodos cervinus", "Diplodus cervinus", fish_filter$species)

# Filter the fish_data data frame to keep only rows where the species is present in unique_species

fish_filter <- fish_data[fish_data$species %in% FB_species, ] %>%
 dplyr:: filter(distance < 5.1)

fish_filter$reserve <- ifelse(fish_filter$site == "Achziv", "R", "N")  
head(fish_filter)

# Filter season
#fish_filter <-   fish_filter%>%
  #filter(season == "Fall")
```

```{r}
fb2use<-FBsize
# Using $ operator with negative indexing
fb2use$Locality <- NULL
fb2use$X <- NULL
fb2use$X.1 <- NULL
head(fb2use)
fb2use <- rename(fb2use, species = Fbspecies_chosen)

fish_filter_a_b <- left_join(fish_filter,fb2use, by= "species")

# Calculate the average size for each species and fill NA values with these averages
fish_filter_a_b <- fish_filter_a_b %>%
  group_by(species) %>%
  mutate(size = ifelse(is.na(size), mean(size, na.rm = TRUE), size)) %>%
  ungroup() %>%
  mutate(fish_weight = a * size^b)

# Display the first few rows
head(fish_filter_a_b)

```
#A. calculate the total number of individuals per species 
norm weight: scale the weight : weight -mean weight devided by the standard deviation

```{r}

fish_long <- fish_filter_a_b %>%
  dplyr::select(-transect, -a_b, -a, -b, -habitat, -depth, -factor_visibility, -observer,  -distance, -visibility, -size, -taxonomy_group) %>%
  arrange(tag)%>% 
  relocate(species, .after = tag) %>%
  dplyr::mutate(
    occ = 1)
#      ) %>% 
#   dplyr::group_by(across(site_id:tag)) %>% 
#   dplyr::mutate(
#        "tag_weight" = sum(fish_weight),  "total_ind" = sum(occ)
#   ) %>% 
#   ungroup()
# head(fish_long)

fish_long_med <- left_join(fish_long,MedSp, by = "species")
  # Summarize fish_weight and occ per tag
fish_summary <- fish_long_med %>%
   relocate(Lessepsian, .after = tag) %>% 
  relocate(depth_category, .after = site_id) %>% 
   relocate(site, .after = season) %>% 
  group_by((across(site:Lessepsian))) %>%
  summarize(
    tag_weight = sum(fish_weight, na.rm = TRUE),  # Summing fish_weight per tag
    total_ind = sum(occ, na.rm = TRUE)            # Summing occ per tag
  )

# View the summary
head(fish_summary)
```
#Overall biomass
```{r}
Biomass <- fish_summary %>%
  filter(site == "Achziv" | site == "SdotYam")
 

Plot_biomass <- Biomass%>%
  ggplot(aes(y = log2(tag_weight /1000))) +
  geom_boxplot(notch=TRUE, notchwidth = 0.8) +
  facet_wrap(~depth_category) +
  theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Log2 of biomass [Kg] per 1000 square m") + theme(axis.text.x = element_blank())
Plot_biomass


```


#Lesspesians
```{r}
Plot_biomassLess <- Biomass %>%
  ggplot(aes(y = log2(tag_weight / 1000), x= site_id, color = Lessepsian)) +  # Use 'color' to partition by Lessepsian
  geom_boxplot(notch = TRUE, notchwidth = 0.3) +
 # facet_wrap(~site_id) +
    ylab("Log2 of biomass [Kg] per 1000 square m") +
theme_classic() +
  scale_color_manual(values = c("0" = "blue", "1" = "red"))  # Reverse color scale: 0 = blue, 1 = red

Plot_biomassLess
```

#Overall abundannce
```{r}
abundannce <- fish_summary %>%
  filter(site == "Achziv" | site == "SdotYam")
 
Plot_abundannce <- abundannce%>%
  ggplot(aes(y = total_ind)) +
  geom_boxplot(notch=TRUE, notchwidth = 0.8) +
  facet_wrap(~site_id) +
  theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Fish abundance per 1000 square m") + theme(axis.text.x = element_blank())
Plot_abundannce


```

### Combining with functional grouping
```{r}
fish_long_Diet <- fish_long_med %>% 
  left_join(Diet, by = "species" )  %>%
  filter(!is.na(Diet))
head(fish_long_Diet)
write.csv(fish_long_Diet, here("R.generated.Data","fish_size_diet.csv" ))
```
### Plotting
#B. Calculate  the total number of species (n_species_sample)and total individuals per survey (n_ind_sample))

```{r}
head(fish_long_Diet)
fish_summaryDiet <- fish_long_Diet %>%
    relocate(Lessepsian, .after = Diet) %>% 
    relocate(site_id, .after = sample_year) %>%
    relocate(season, .after = sample_year) %>% 
  relocate(site, .after = site_id) %>% 
  relocate(Diet, .after = date_id) %>%
      relocate(depth_category, .after = Diet) %>% 
  relocate(tag, .after = depth_category) %>% 
  group_by((across(site:depth_category))) %>%
  summarize(
    tag_weight = sum(fish_weight, na.rm = TRUE),  # Summing fish_weight per tag
    total_ind = sum(occ, na.rm = TRUE)            # Summing occ per tag
  )

# View the summary
head(fish_summaryDiet)

```

```{r}
fish_wide<- spread(fish_summaryDiet,Diet, tag_weight,fill = 0)
head(fish_wide)
colnames(fish_wide)

# df <-fish_summaryDiet%>%
#   filter(site %in% c("Achziv", "SdotYam")) %>%
#   group_by(depth_category) %>%
#   wilcox_test(visibility ~ site) %>%
#   adjust_pvalue(method = "bonferroni")
# df
```
#Diet  Biomass Color scale
```{r}

colorDiet = c("FC" = "red", "H" = "green4", "Mi" = "darkorange","Ma" = "brown","O" = "grey", "PK"= "darkblue")
fish_DietP<-fish_summaryDiet %>%
  filter(site == "Achziv" | site == "SdotYam" ) %>%
  group_by(depth_category, Diet) %>%
  summarize(avg_biomass = mean(tag_weight, na.rm = TRUE))
# Create the stacked bar plot
ggplot(data = fish_DietP, aes(x = as.factor(depth_category), y = avg_biomass/1000, fill = Diet)) +
  geom_col( position = "stack") +
  scale_fill_manual(values = colorDiet) +  # Apply the custom color palette
  theme_classic() +
  labs(
    x = "Site ID",
    y = "Log of Mean Fish Biomass per 1000 sqm",
    fill = "Diet"
  ) +
  theme(
    text = element_text(size = 16),  # Set overall text size to 18
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
  )+scale_y_log10()
```
```{r}
ggplot(data = fish_DietP, aes(x = as.factor(depth_category), y = avg_biomass/1000, fill = Diet)) +
  geom_col( position = "stack") +
  scale_fill_manual(values = colorDiet) +  # Apply the custom color palette
  theme_classic() +
  labs(
    x = "Site ID",
    y = "Mean Fish Biomass[Kg] per 1000 sqm",
    fill = "Diet"
  ) +
  theme(
    text = element_text(size = 16),  # Set overall text size to 18
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
  )
```
#Abundance

```{r}

colorDiet = c("FC" = "red", "H" = "green4", "Mi" = "darkorange","Ma" = "brown","O" = "grey", "PK"= "darkblue")
fish_DietAbundP<-fish_summaryDiet %>%
  filter(site == "Achziv" | site == "SdotYam" ) %>%
  group_by(depth_category, Diet) %>%
  summarize(avg_abund = mean(total_ind, na.rm = TRUE))
# Create the stacked bar plot
ggplot(data = fish_DietAbundP, aes(x = as.factor(depth_category), y = avg_abund, fill = Diet)) +
  geom_col( position = "stack") +
  scale_fill_manual(values = colorDiet) +  # Apply the custom color palette
  theme_classic() +
  labs(
    x = "Site ID",
    y = "Log of Mean Abundnace per 1000 sqm",
    fill = "Diet"
  ) +
  theme(
    text = element_text(size = 16),  # Set overall text size to 18
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
  )+scale_y_log10()
```
```{r}
ggplot(data = fish_DietP, aes(x = as.factor(depth_category), y = avg_biomass/1000, fill = Diet)) +
  geom_col( position = "stack") +
  scale_fill_manual(values = colorDiet) +  # Apply the custom color palette
  theme_classic() +
  labs(
    x = "Site ID",
    y = "Mean Fish Biomass[Kg] per 1000 sqm",
    fill = "Diet"
  ) +
  theme(
    text = element_text(size = 16),  # Set overall text size to 18
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
  )
```
#The effect of depth_category on Diet
```{r}

fish_summaryH<-fish_summaryDiet %>%
  filter(site == "Achziv" |"SdotYam" ) %>%
  filter(Diet == "H") 

 fish_plotH <-ggplot(fish_summaryH, aes(x=total_ind,
           y=-depth_category,
           color = site))+
        xlab("Herbivore abundance")+
       ylab("Depth category")+  geom_jitter() +theme_classic()  #+      ggtitle("Relationship between Trophic group abundance and biomass")+
fish_plotH
scale.lm=lm(total_ind~depth_category, data=fish_summaryH) #simple linear model for extracting the slope
observed_slp=scale.lm$coef[2] #extracting the value of the slope from the model
observed_slp #prints the value of the slope
summary(scale.lm)
```
#Randomization testing 
```{r}
rand_slopes = replicate(1000,{
  random_x = sample(fish_summaryH$depth_category,length(fish_summaryH$depth_category),replace = F) #sample with no replacements 
  random_model = lm(fish_summaryH$total_ind ~ random_x) 
  random_model$coef[2]
})
```
```{r}
hist(rand_slopes,xlim=c(-1.5,1.5))
abline(v=observed_slp, col='red')
```
#Obtaining p value
```{r}

sum( rand_slopes<=  observed_slp)

# Or in terms of a p-value:

sum( rand_slopes<=  observed_slp)/1000
```



```{r}

fish_summaryDiet %>%
  filter(site == "Achziv" | site == "SdotYam") %>%
  ggplot(aes(x=log10(total_ind),
           y=log2(tag_weight/1000),
           color = Diet))+
        xlab("Log10(Trophic abundance per sampling day)")+
       ylab("Log2(Biomass, Kg)")+
      ggtitle("Relationship between Trophic group abundance and biomass")+
  geom_jitter()+
    facet_wrap(~site_id ) +theme_classic()  +
  scale_color_manual(values = colorDiet) # Specify color palette
```

##plotting the relationship between species and individuals number:
```{r}
colorDiet = c("red", "green4", "darkorange","brown","grey", "darkblue")
fish_summaryDiet %>%
  filter(site == "Achziv" | site == "SdotYam") %>%
  ggplot(aes(x=log10(total_ind),
           y=log2(tag_weight/1000),
           color = Diet, shape = Lessepsian))+
        xlab("Log10(Trophic abundance per sampling day)")+
       ylab("Log2(Biomass, Kg)")+
      ggtitle("Relationship between Trophic group abundance and biomass")+
  geom_jitter()+
    facet_wrap(~site_id ) +theme_classic()  +
  scale_color_manual(values = colorDiet) # Specify color palette
```
#Lesspesian Abundance
```{r}
Plot_abundLess <- Biomass %>%
  ggplot(aes(y = total_ind, x = site_id, color = Lessepsian)) +  # Use 'color' to partition by Lessepsian and treat it as a factor
  geom_boxplot(notch = TRUE, notchwidth = 0.3) +
  ylab("Fish abundance per 1000 square m") +
  theme_classic() +
  scale_color_manual(values = c("0" = "blue", "1" = "red"))  # Reverse color scale: 0 = blue, 1 = red

Plot_abundLess

```
