---
title: "Data processing for chapter 3: 25-45 m sites" MKMRS data of fish survey  and invertebrate photoquadrates
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
WHEN UPDATING A NEW FISH_SURVEY FILE- CORRECT VIVIBILITY TO 7 AT AchzivR453Spring2020
AchzivR454Spring2020

---

### step 1 load data library etc and set wd


```{r setup, include=FALSE}



library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
library(assertthat)
library(devtools)
#For Diversity calculations
library(plotrix)
library(betapart)
library(here)
library(ggridges)

library(ggtext)# for italics in the plot legend 
library(ggpubr) 
# Richness
library(rareNMtests)
library(mobr)
here<-here::here


```

```{r}
# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")
csv_path2 <- here("data", "45FBsize_a_b_with_missing.csv")
# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
FBsize <- read.csv(csv_path2)
complex.inic <-read.csv(here("data", "EnvVar10_45.csv"))
Diet <- read_csv(here("data","Diet.csv"))
fish_data <-merge(fish_data,sites, by= "site_id") 
trait <-read.csv(here("data", "Belmaker_Species_Traits.csv"))

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))
#Make sure Achziv is Reserve and Sdot Yam Not
head(trait)
MedSp <-select(trait,"Species_updated_name_fishbase", "Lessepsian")

head(fish_data)
colnames(MedSp)<-c("species", "Lessepsian")
MedSp$Lessepsian <-   as.factor(MedSp$Lessepsian)


```


###QA
```{r}

# Group by "tag" and count unique "sample_date" values
result <- fish_data %>%
  group_by(tag) %>%
  summarise(unique_dates = n_distinct(sample_date)) %>%
  filter(unique_dates > 1)

# Print the result
print(result)
```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs
Filtering species with Fishbase biomass constants
###Renaming fish name spelling errors; renaming reserve status; cropping distance from transect 
```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter<-fish_data %>%dplyr:: filter((site=="Achziv" | site=="SdotYam"))
# Get the unique species names from the FBsize data frame
FB_species <- unique(FBsize$Fbspecies_chosen)
fish_filter$species <- sub("Salpa salpa", "Sarpa salpa", fish_filter$species)
fish_filter$species <- sub("Saprisoma cretense", "Sparisoma cretense", fish_filter$species)
fish_filter$species <- sub("Diplodos cervinus", "Diplodus cervinus", fish_filter$species)

# Filter the fish_data data frame to keep only rows where the species is present in unique_species

fish_filter <- fish_data[fish_data$species %in% FB_species, ] %>%
 dplyr:: filter(distance < 5.1)

fish_filter$reserve <- ifelse(fish_filter$site == "Achziv", "R", "N")  
head(fish_filter)

# Filter season
#fish_filter <-   fish_filter%>%
  #filter(season == "Fall")
```

```{r}
fb2use<-FBsize
# Using $ operator with negative indexing
fb2use$Locality <- NULL
fb2use$X <- NULL
fb2use$X.1 <- NULL
head(fb2use)
fb2use <- rename(fb2use, species = Fbspecies_chosen)

fish_filter_a_b <- left_join(fish_filter,fb2use, by= "species")
fish_filter_a_b <- fish_filter_a_b %>%
  mutate(fish_weight =  a * size^b )
head(fish_filter_a_b)

```
#A. calculate the total number of individuals per species 
norm weight: scale the weight : weight -mean weight devided by the standard deviation

```{r}

fish_long <- fish_filter_a_b %>%
  dplyr::select(-transect, -a_b, -a, -b, -habitat, -depth, -factor_visibility, -observer, -size, -distance, -tag, -taxonomy_group) %>%
  dplyr::relocate(species, .after = lat) %>%
  dplyr::mutate(
    occ = 1,
    norm_wt = scale(fish_weight , center = TRUE, scale = TRUE)
  ) %>% 
  dplyr::group_by(across(site_id:species)) %>% 
  dplyr::summarise(
    total_ind = sum(occ),
    sp_weight = sum(fish_weight),
    Av.norm = mean(norm_wt)
  ) %>% 
  ungroup()
fish_long <- left_join(fish_long,MedSp, by = "species")
  
head(fish_long)
```
#B. Calculate  the total number of species (n_species_sample)and total individuals per survey (n_ind_sample))

```{r}
fish_long<-fish_long %>%
  dplyr::group_by(across(site_id:species)) %>% 
  dplyr::mutate(n_species_sample = n_distinct(species),
          n_ind_sample= sum(total_ind)) %>%
  dplyr::ungroup()
head(fish_long)
```
### Combining with functional grouping
```{r}
fish_long_Diet <- fish_long %>% 
  left_join(Diet, by = "species" )  %>%
  filter(!is.na(Diet))
head(fish_long_Diet)
write.csv(fish_long_Diet, here("R.generated.Data","fish_size_diet.csv" ))
```
### Plotting

##plotting the relationship between species and individuals number:
```{r}
colorDiet = c("red", "green4", "darkorange","brown","grey", "darkblue")
fish_long_Diet %>%
  filter(site == "Achziv" | site == "SdotYam") %>%
  ggplot(aes(x=log10(n_ind_sample),
           y=log2(sp_weight/1000),
           color = Diet))+
        xlab("Log10 of abundance per 1,000 sqare m")+
       ylab("Log2 of Weight, Kg")+
      ggtitle("Relationship between Trophic group abundance and weight")+
  geom_jitter()+
    facet_wrap(~site_id ) +theme_classic()  +
  scale_color_manual(values = colorDiet) # Specify color palette
```
```{r}
color6 = c("lightblue3", "blue2", "darkblue","yellow3", "darkorange", "brown")
Biomass <- fish_long_Diet %>%
  filter(site == "Achziv" | site == "SdotYam")%>%
  dplyr::relocate(Diet, .after = lat) %>%
  dplyr::relocate(Lessepsian, .after = date_id) %>% 
  dplyr::group_by(across(site_id:lat)) %>% 
  dplyr::summarise(Diet_biomass = sum(sp_weight)) 
Plot_biomass <- Biomass%>%
  ggplot(aes(y = log2(Diet_biomass /1000))) +
  geom_boxplot(notch=TRUE, notchwidth = 0.8) +
  facet_wrap(~site_id) +
  theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Log2 of biomass [Kg] per 1000 square m") + theme(axis.text.x = element_blank())
#Lessepsian biomass
Plot_biomass <- Biomass %>%
  ggplot(aes(y = log2(Diet_biomass / 1000), fill = Lessepsian)) +  # Use 'fill' for stacked effect or 'color' for borders
  geom_boxplot(notch = TRUE, notchwidth = 0.8) +
 
  theme_classic() +
  scale_fill_manual(values = c(1 = "red", 0 = "blue")) +  # Customize colors
  ylab("Log2 of biomass [Kg] per 1000 square m") +
  theme(axis.text.x = element_blank())
# facet_wrap(~site_id) +
print(Plot_biomass)

```

```{r}
# Define breaks for the histogram with each bin twice the value of the preceding one
custom_breaks <- 2^(0:10)
fish_wide %>%
   ggplot(aes( x= log2(n_ind_sample),   y = n_species_sample, color= season))+ geom_col(breaks = custom_breaks)+
        xlab("log2(Abundance) at 10 m depth sites")+
       ylab("No. of species")+ theme_classic() + scale_y_continuous(limits = c(0, 30))
```
```{r}
# Example bxpT
bxpT <- ggplot(my_data, aes(x = site_id, y = LogVR)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Log(Vertical Relief)")

# Set y-axis limits without truncating data points
bxpT <- bxpT + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpT)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpT <- bxpT     + labs(x = NULL)   # 
bxpT
```


```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpS <- ggplot(my_data, aes(x = site_id, y = Rugosity)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Rugosity")

# Set y-axis limits without truncating data points
#bxpS <- bxpS + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpS)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpS <- bxpS     + labs(x = NULL)   # 
bxpS
```


```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpH <- ggplot(my_data, aes(x = site_id, y = Slope)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Slope")

# Set y-axis limits without truncating data points
#bxpH <- bxpH + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpH)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpH <- bxpH     + labs(x = NULL)   # 
bxpH
```
```{r}
# Assuming you have a ggplot object named bxpT
# Replace it with your actual ggplot object

# Example bxpT
bxpC <- ggplot(my_data, aes(x = site_id, y = Curvature)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Site id", y = "Curvature")

# Set y-axis limits without truncating data points
#bxpC <- bxpC + scale_y_continuous(limits = c(0, 1))

# Print the modified ggplot
print(bxpC)

#scale_y_continuous(expand = c(0, 0)) 
# Remove x-axis title
bxpC <- bxpC     + labs(x = NULL)   # 
bxpC
```
### ANOVA testing
```{r}
# Convert depth_category to factor for correct analysis
my_data$depth_category <- as.factor(my_data$depth_category)
# my_data_sy <-my_data %>%
#   filter(site == "SdotYam")


# Perform two-way ANOVA for SY
anova_result <- aov(LogVR ~ site * depth_category , data = my_data)
summary(anova_result)
```


```{r}
library(ggpubr)
# Set a common x-axis titlec ommon.legend = TRUE,  common.xlabel = "N" )
tiff("Habitat heterogeneity indices.jpeg", units="in", width=7, height=6, res=600)
diversPlots <- ggarrange( bxpT, bxpS, bxpH, bxpC, 
          labels = c("       a ", "      b ", "       c ", "      d   "),
          ncol = 2, nrow = 2)   # Set a common x-axis title)
diversPlots
dev.off()
diversPlots

```

```{r}
# Set up a 1 row, 5 column layout for plotting
par(mfrow=c(1, 6))

# Loop through each Diet group and create a boxplot
for (i in 1:min(6, length(diet_groups))) {
  if (i <= length(diet_groups)) {
    boxplot(norm_wt ~ site_id, data = diet_groups[[i]],
            main = names(diet_groups)[i],
            xlab = "NULL", ylab = "NULL")
  } else {
    plot.new() # Create an empty plot for remaining panels
  }
}


```

```{r}
# Split the data frame by Diet
add_sample <- function(x){
   return(c(y = max(x) + .025, 
            label = length(x)))
}
diet_groups <- split(fish_filter_a_b, fish_filter_a_b$Diet)


# Define your color vector
color6 <- c("lightblue3", "blue2", "darkblue", "yellow3", "darkorange", "brown")

# Loop through each Diet group and create a boxplot
for (i in 1:min(6, length(diet_groups))) {
  plot_name <- paste0("size", i, ".png")
  
  if (i <= length(diet_groups)) {
    png(plot_name, width = 800, height = 600) # Open a PNG graphics device
    boxplot(norm_wt ~ site_id, data = diet_groups[[i]],
            main = names(diet_groups)[i],
            xlab = "Site ID", ylab = "Normalized Weight",
            col = color6[1:length(unique(diet_groups[[i]]$site_id))],
            border = color6[1:length(unique(diet_groups[[i]]$site_id))],
            pch = 21, bg = color6[1:length(unique(diet_groups[[i]]$site_id))])
    # Add labels for number of samples
    add_sample_values <- add_sample(diet_groups[[i]]$norm_wt)
    text(x = 1, y = add_sample_values["y"], labels = add_sample_values["label"], pos = 1, cex = 0.8, col = "black")
    
    
    dev.off() # Close the graphics device to save the plot
  } else {
    plot.new() # Create an empty plot for remaining panels
  }
}


```


```{r}
library(ggrepel)

# Split the data frame by Diet
add_sample <- function(x){
  return(c(y = max(x) + .025, 
           label = length(x)))
}
diet_groups <- split(fish_filter_a_b, fish_filter_a_b$Diet)

# Define your color vector
color6 <- c("lightblue3", "blue2", "darkblue", "yellow3", "darkorange", "brown")

# Loop through each Diet group and create a boxplot
for (i in 1:min(5, length(diet_groups))) {
  plot_name <- paste0("size", i, ".png")
  
  if (i <= length(diet_groups)) {
    png(plot_name, width = 800, height = 600) # Open a PNG graphics device
    
    # Create a ggplot object
    p <- ggplot(diet_groups[[i]], aes(x = as.factor(site_id), y = norm_wt, fill = as.factor(site_id))) +
      geom_boxplot() +
      labs(title = names(diet_groups)[i], x = "Site ID", y = "Normalized Weight") +
      scale_fill_manual(values = color6[1:length(unique(diet_groups[[i]]$site_id))]) +
      theme_minimal()+
  stat_summary(
    geom = "text",
    fun.data = add_sample,
    aes(label = paste("n =", ..label..),
        color = species,
        color = after_scale(darken(color, .1, space = "HLS"))),
    family = "Roboto Condensed",
    size = 4,
   position = position_dodge(width = 0.9)
  )  
      
    
    print(p)
    dev.off() # Close the graphics device to save the plot
  } else {
    plot.new() # Create an empty plot for remaining panels
  }
}

    
```
