---
title: "Data processing for chapter 3: 25-45 m sites" MKMRS data of fish survey  and invertebrate photoquadrates
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
WHEN UPDATING A NEW FISH_SURVEY FILE- CORRECT VIVIBILITY TO 7 AT AchzivR453Spring2020
AchzivR454Spring2020

---

### step 1 load data library etc and set wd


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
library(assertthat)
library(devtools)
#For Diversity calculations
library(plotrix)
library(betapart)
library(here)
library(ggridges)

library(ggtext)# for italics in the plot legend 
library(ggpubr) 
# Richness
library(rareNMtests)
library(mobr)
here<-here::here


```

```{r}
# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")
csv_path2 <- here("data", "45FBsize_a_b_with_missing.csv")
# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
FBsize <- read.csv(csv_path2)
complex.inic <-read.csv(here("data", "EnvVar10_45.csv"))
Diet <- read_csv(here("data","Diet.csv"))
fish_data <-merge(fish_data,sites, by= "site_id") 

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))
#Make sure Achziv is Reserve and Sdot Yam Not

head(fish_data)
```

###QA
```{r}

# Group by "tag" and count unique "sample_date" values
result <- fish_data %>%
  group_by(tag) %>%
  summarise(unique_dates = n_distinct(sample_date)) %>%
  filter(unique_dates > 1)

# Print the result
print(result)
```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs
Filtering species with Fishbase biomass constants
###Renaming fish name spelling errors; renaming reserve status; cropping distance from transect 
```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter<-fish_data %>%dplyr:: filter((site=="Achziv" | site=="SdotYam"))
# Get the unique species names from the FBsize data frame
FB_species <- unique(FBsize$Fbspecies_chosen)
fish_filter$species <- sub("Salpa salpa", "Sarpa salpa", fish_filter$species)
fish_filter$species <- sub("Saprisoma cretense", "Sparisoma cretense", fish_filter$species)
fish_filter$species <- sub("Diplodos cervinus", "Diplodus cervinus", fish_filter$species)

# Filter the fish_data data frame to keep only rows where the species is present in unique_species

fish_filter <- fish_data[fish_data$species %in% FB_species, ] %>%
 dplyr:: filter(distance < 5.1)

fish_filter$reserve <- ifelse(fish_filter$site == "Achziv", "R", (ifelse(fish_filter$site_id == "SdotYam45", "N", "NR"))) # if its true write R if its false write Fall
head(fish_filter)

# Filter season
#fish_filter <-   fish_filter%>%
  #filter(season == "Fall")
```

```{r}
fb2use<-FBsize
# Using $ operator with negative indexing
fb2use$Locality <- NULL
fb2use$X <- NULL
fb2use$X.1 <- NULL
head(fb2use)
fb2use <- rename(fb2use, species = Fbspecies_chosen)

fish_filter_a_b <- left_join(fish_filter,fb2use, by= "species")
fish_filter_a_b <- fish_filter_a_b %>%
  mutate(fish_weight =  a * size^b )
head(fish_filter_a_b)
fish_filter_a_b <- fish_filter_a_b %>%
  dplyr::group_by(species) %>%
  dplyr::mutate(norm_wt = (fish_weight- min(fish_weight, na.rm = TRUE)) / (max(fish_weight, na.rm = TRUE) - min(fish_weight, na.rm = TRUE)))%>%
  ungroup() 

head(fish_filter_a_b)
# Subset the data frame for Siganus rivulatus
siganus_data <- subset(fish_filter_a_b, species == "Siganus rivulatus")

# Summarize the fish_weight column
summary(siganus_data$fish_weight)
```
### Combining with functional grouping
```{r}
fish_filter_a_b <- fish_filter_a_b %>% 
  left_join(Diet, by = "species" ) 
head(fish_filter_a_b)
write.csv(fish_filter_a_b, here("R.generated.Data","fish_filter_a_b.csv" ))
```
### Plotting

```{r}


# Split the data frame by Diet
# Filter out sites that are not "SdotYam" or "Achziv"
fish_filter_a_b <- fish_filter_a_b %>%
  dplyr::filter(site %in% c("SdotYam", "Achziv")) %>% 
  mutate(Year_season_id =paste0(sample_year,season,site_id)) %>%
  relocate("Year_season_id",.after = tag)   %>% 
  dplyr::select(site,date_id, Year_season_id, site_id, depth_category, season, species,norm_wt, Diet, fish_weight) %>% 
  dplyr::group_by(site,date_id, Year_season_id, site_id, depth_category, season, species, Diet, fish_weight,species()) %>% # use this unit for analyses
  dplyr::summarise(total_wt_sp = sum(fish_weight, na.rm = TRUE)) 
head(fish_filter_a_b)
# Split the data frame by Diet
diet_groups <- split(fish_filter_a_b, fish_filter_a_b$Diet)

# Set up a 1 row, 5 column layout for plotting
par(mfrow=c(1, 6))

# Loop through each Diet group and create a boxplot
for (i in 1:min(6, length(diet_groups))) {
  if (i <= length(diet_groups)) {
    boxplot(norm_wt ~ site_id, data = diet_groups[[i]],
            main = names(diet_groups)[i],
            xlab = "NULL", ylab = "NULL")
  } else {
    plot.new() # Create an empty plot for remaining panels
  }
}


```

```{r}
# Split the data frame by Diet
add_sample <- function(x){
   return(c(y = max(x) + .025, 
            label = length(x)))
}
diet_groups <- split(fish_filter_a_b, fish_filter_a_b$Diet)


# Define your color vector
color6 <- c("lightblue3", "blue2", "darkblue", "yellow3", "darkorange", "brown")

# Loop through each Diet group and create a boxplot
for (i in 1:min(6, length(diet_groups))) {
  plot_name <- paste0("size", i, ".png")
  
  if (i <= length(diet_groups)) {
    png(plot_name, width = 800, height = 600) # Open a PNG graphics device
    boxplot(norm_wt ~ site_id, data = diet_groups[[i]],
            main = names(diet_groups)[i],
            xlab = "Site ID", ylab = "Normalized Weight",
            col = color6[1:length(unique(diet_groups[[i]]$site_id))],
            border = color6[1:length(unique(diet_groups[[i]]$site_id))],
            pch = 21, bg = color6[1:length(unique(diet_groups[[i]]$site_id))])
    # Add labels for number of samples
    add_sample_values <- add_sample(diet_groups[[i]]$norm_wt)
    text(x = 1, y = add_sample_values["y"], labels = add_sample_values["label"], pos = 1, cex = 0.8, col = "black")
    
    
    dev.off() # Close the graphics device to save the plot
  } else {
    plot.new() # Create an empty plot for remaining panels
  }
}


```


```{r}
library(ggrepel)

# Split the data frame by Diet
add_sample <- function(x){
  return(c(y = max(x) + .025, 
           label = length(x)))
}
diet_groups <- split(fish_filter_a_b, fish_filter_a_b$Diet)

# Define your color vector
color6 <- c("lightblue3", "blue2", "darkblue", "yellow3", "darkorange", "brown")

# Loop through each Diet group and create a boxplot
for (i in 1:min(5, length(diet_groups))) {
  plot_name <- paste0("size", i, ".png")
  
  if (i <= length(diet_groups)) {
    png(plot_name, width = 800, height = 600) # Open a PNG graphics device
    
    # Create a ggplot object
    p <- ggplot(diet_groups[[i]], aes(x = as.factor(site_id), y = norm_wt, fill = as.factor(site_id))) +
      geom_boxplot() +
      labs(title = names(diet_groups)[i], x = "Site ID", y = "Normalized Weight") +
      scale_fill_manual(values = color6[1:length(unique(diet_groups[[i]]$site_id))]) +
      theme_minimal()+
  stat_summary(
    geom = "text",
    fun.data = add_sample,
    aes(label = paste("n =", ..label..),
        color = species,
        color = after_scale(darken(color, .1, space = "HLS"))),
    family = "Roboto Condensed",
    size = 4,
   position = position_dodge(width = 0.9)
  )  
      
    
    print(p)
    dev.off() # Close the graphics device to save the plot
  } else {
    plot.new() # Create an empty plot for remaining panels
  }
}

    
```



##before we will convert the data to the wide format and start plotting we will add two more summary columns that will come to use later on. lets calculate the total number of individuals per survey and the total number of species.
#A. calculate the total number of individuals per species 
```{r}

fish_long <- fish_filter_a_b %>%
  dplyr::select(-transect,-a_b,-habitat, -depth, -factor_visibility, -observer,-distance, -size, -a, -b) %>%
  dplyr::relocate(species,.after = tag)  %>% 
 dplyr::group_by(across(site_id:species)) %>% 
 dplyr::summarise(sp_weight = sum(fish_weight))
head(fish_long)
```
#B. Calculate  the total number of species (n_species_sample)and total biomass per survey (n_ind_sample))

```{r}
fish_long<-fish_long %>%
  dplyr::group_by(tag)%>% 
  dplyr::mutate(n_species_sample = n_distinct(species),
          n_ind_sample= sum(sp_weight)) %>%
  dplyr::ungroup()
head(fish_long)




```

###Finally, lets convert our data to wide format:


```{r}
fish_wide<- spread(fish_long,species,sp_weight,fill = 0)
head(fish_wide)
colnames(fish_wide)


```

###making Fish abudance table filtering empty species

```{r}


fish_wide <- fish_wide[, colSums(fish_wide != 0) > 0]

write.csv(fish_wide, here("R.generated.Data", "45Fish_data_clean.csv"))

```

####Diversity
```{r}
my_data <- fish_wide %>%
  mutate(Year_season_id =paste0(sample_year,season,site_id)) %>%
  relocate("Year_season_id",.after = tag) 
 head(my_data)
 colnames(my_data)
```

###Species Matrix +Metadata 
```{r}
colnames(my_data)
first_species <- 17
sp_matrix <- my_data[,first_species:ncol(my_data)] %>%
  select_if(~ sum(.) != 0)
meta_data <- my_data[,1:16] # columns i want to keep from my data


```
### Hill numbers profile Allways use  hill = T
```{r}
renyi_profile <- renyi(sp_matrix,  
                       scales = c(0, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 4, 8, 16, 32, 64, Inf),
                       hill = T)


renyi_df<-bind_cols(meta_data,renyi_profile)
head(renyi_df)
renyi_df<-gather(renyi_df,"Q","Value",17:31)
head(renyi_df)
renyi_df<-renyi_df %>% arrange(site_id,depth_category,sample_year, season,tag) %>%
   filter(!is.na(Q))

renyi_df$Q <- as.numeric(renyi_df$Q)
```

```{r}
head(renyi_df)

color6 = c("lightblue3", "blue2", "darkblue","yellow3", "darkorange", "brown")
ggplot(data = renyi_df) +
  aes(x = Q, y = Value, group = tag, color = factor(site_id)) +
  geom_line() +
  scale_x_continuous(limits = c(0, 10)) +
  scale_y_continuous(limits = c(0, 13)) +
  facet_wrap(. ~ site_id) +
  labs(y = expression("Diversity  D(p) "^"Q")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  )
#  scale_color_manual(values = color6)  # Specify color palette



```

###SEction 2: Grouping per sampling day
```{r}
colnames(my_data)
group_year_data <- my_data %>% 
  dplyr::select(date_id, Year_season_id, site_id, depth_category, season, "Acantholabrus palloni":"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni":"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_year_data)
colnames(group_year_data)

colnames(log10_hel_my_data)
group_trans_data <- log10_hel_my_data %>% 
  dplyr::select(date_id, site, Year_season_id, site_id, depth_category, season, "Acantholabrus palloni":"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id, site, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni":"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_trans_data)



```


```{r}

renyi_profile_group <- renyi(group_year_data[,-c(1:5)],
                             scales = c(0, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 4, 8, 16, 32, 64, Inf),  hill = T)

renyi_df_group<-bind_cols(group_year_data[,1:5],renyi_profile_group)

#compose long df with  new Q grouping
renyi_df_group<-gather(renyi_df_group,"Q","Value",6:20)

renyi_df_group<-renyi_df_group %>% arrange(Year_season_id, site_id,depth_category, season )

renyi_df_group$Q <- as.numeric(renyi_df_group$Q)
renyi_df_group$depth_category <- as.factor(renyi_df_group$depth_category)
head(renyi_df_group)
```
```{r}
ggplot(data = renyi_df_group) +
  aes(x = Q, y = Value, group = Year_season_id, color = factor(site_id)) +
  geom_line() +
  scale_x_continuous(limits = c(0, 10)) +
  scale_y_continuous(limits = c(0, 20)) +
  facet_wrap(. ~ site_id) +
  labs(y = expression("Diversity  D(p) "^"Q")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) +
  scale_color_manual(values = color6)  # Specify color palette
```
###creating unique lists of species
###Now with transect transformed data: log10_hel_my_data 
```{r}



bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(my_data$site_id)){
  
site_data <- my_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

NTTransectBeta <- ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6) # Specify color palette

NTTransectBeta
```

###Now with transect transformed data: log10_hel_my_data 
```{r}



bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(log10_hel_my_data$site_id)){
  
site_data <- log10_hel_my_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

TransectBeta <- ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6) # Specify color palette

TransectBeta
```
# grouped data

```{r}



bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(group_year_data$site_id)){
  
site_data <- group_year_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

pGr <-ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6)  # Specify color palette

pGr


```
###Now with transformed data: log10_hel_my_data grouped in group_trans_data 

```{r}
bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(group_trans_data$site_id)){
  
site_data <- group_trans_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame
```


```{r}
# lets plot:

PGrTrans <- ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6)  # Specify color palette

PGrTrans
NTTransectBeta <- NTTransectBeta + scale_y_continuous(limits = c(0.2, 0.8),  
                               breaks = seq(0.2, 0.8, by = 0.2)) + theme(axis.title.x = element_blank(),  # Remove x-axis title
                 axis.text.x = element_blank()) +
           labs(y = "Bray-Curtis dissimilarity")

TransectBeta <- TransectBeta + scale_y_continuous(limits = c(0.2, 0.8),  
                               breaks = seq(0.2, 0.8, by = 0.2)) +   theme(axis.title.x = element_blank(),  
                                                                           axis.title.y = element_blank()  ,# Remove x-axis title
                 axis.text.x = element_blank()) +
           labs(y = "Bray-Curtis dissimilarity")

pGr <- pGr + scale_y_continuous(limits = c(0.2, 0.8),  
                               breaks = seq(0.2, 0.8, by = 0.2)) + theme(axis.title.x = element_blank(),  # Remove x-axis title
                 axis.text.x = element_blank()) +
           labs(y = "Bray-Curtis dissimilarity")

PGrTrans <- PGrTrans + scale_y_continuous(limits = c(0.2, 0.8),  
                               breaks = seq(0.2, 0.8, by = 0.2)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()  , # Remove x-axis title
                 axis.text.x = element_blank()) +
           labs(y = "Bray-Curtis dissimilarity")
plot.new()

TurnBetaPlots <- ggarrange( NTTransectBeta, TransectBeta,pGr, PGrTrans, 
          labels = c("  Transect ", "Transformed Transect", "  Grouped", "Transformed grouped"),          ncol = 2, nrow = 2, common.legend = TRUE )  
     
 
 
TurnBetaPlots



```

```{r}
site_id_distribution <- table(fish_wide$site_id)
print(site_id_distribution)
```
```{r}
least_sampled <- 24

least_sampled
```
##Incidence data- beta pairs
```{r}
AC_incidence <- sp_matrix %>% 
  replace(sp_matrix > 0, 1) 
AC_incidence_bind<-bind_cols(meta_data,AC_incidence )


caves_beta_pairs <- beta.pair(AC_incidence)
```


```{r message=FALSE, results='hide'}


caves_beta_sample <- beta.sample(AC_incidence, 
            sites= least_sampled,  # how many knolls to randomly sample
            samples = 10) # how many repetitions of (10) knolls to do 


```


after we understood how the function is working lets calculate the multi-site beta diversity of all the years for the site_id   
###Sorensen beta diversity: per transect and grouped

```{r message=FALSE, warning=FALSE, results='hide'}


caves <-my_data  

caves[first_species:ncol(caves)] <-replace(caves[first_species:ncol(caves)],
                                            caves[first_species:ncol(caves)] > 0,
                                            1) # incident data

caves_all_years<-list() # empty list

for (i in unique(caves$site_id)) { # loop
  
  one_year <-  caves %>% filter(site_id == i)
  
  one_year_sp_matrix <- one_year[,first_species:ncol(one_year)]
  
  caves_beta_sample <- beta.sample(one_year_sp_matrix,
            sites= 5, 
            samples = 24)
  
 
  caves_beta_sample_data <- data.frame(mean = caves_beta_sample[["mean.values"]],
                                    se = caves_beta_sample[["sd.values"]])

  caves_beta_sample_data <-rownames_to_column(caves_beta_sample_data,'Component') 

  caves_beta_sample_data$Year_season <- i
  
 caves_all_years[[i]]<-caves_beta_sample_data
  
}


caves_all_years<-bind_rows(caves_all_years)


```


```{r}
caves_all_years<-caves_all_years %>% filter(Component != "beta.SOR")

ggplot(caves_all_years)+
  aes(x=Year_season,y=mean,fill=Component)+
  geom_bar(stat="identity",position = position_stack())+
  ylab(expression(paste(beta,"-diversity" (Sorensen)))) +theme_classic() +
  xlab("Site id")




```
We will use the function `rarefaction.sample` from the package `rareNMtests`. 

In this function the first argument is the data name,  the method argument `sample-size` means that the function will use incidence data and `q = 0` is used for species richness rarefaction.  

Now let's create sample based rarefaction curve for Fall 2018.  
#  Sample based  rarefraction and individual based 

```{r}
sample_based_rare<-list()

for (i in unique(my_data$site_id)) {

one_year<-my_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,first_species:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q = 0)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Richness","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

Richp <- ggplot(sample_based_rare,aes(x=Samples,y= Richness,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 

Richp
```
#Simpson
```{r}
sample_based_rare<-list()

for (i in unique(my_data$site_id)) {

one_year<-my_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,first_species:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q = 2)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","True.diversity","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

TrueDp <- ggplot(sample_based_rare,aes(x=Samples,y= True.diversity,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
TrueDp

```


```{r}
sample_based_rare<-list()

for (i in unique(my_data$site_id)) {

one_year<-my_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,first_species:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q = 1)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Effective.diversity","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

EffDp <- ggplot(sample_based_rare,aes(x=Samples,y= Effective.diversity,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
EffDp

```



```{r}
RarePlots <- ggarrange( Richp, EffDp,TrueDp,
          labels = c("Richness ", "Effective diveristy", "True diversity"),          ncol = 3, nrow = 1, common.legend = TRUE )  # Set a common x-axis title)
     
 RarePlots

```

###individual based rarefaction

```{r message=FALSE, warning=FALSE}

ind_based_rare_all<-list() # empty list

for (i in 1:nrow(group_year_data)) { # for each row...

one_knoll <- group_year_data[i,] # filter
one_knoll_sp_matrix <- one_knoll[,6:ncol(one_knoll)] # create sp_matrix 6 in the first species

rarefaction <- rarefaction.individual(one_knoll_sp_matrix, method = "sample-size", q = 0) # apply rarefaction function

rarefaction$Year_season<-one_knoll$Year_season_id # add which year_season to the data
rarefaction$survey_id<-one_knoll$site_id # add the survey_id to the data
ind_based_rare_all[[i]]<- rarefaction # save in my list
  
}

ind_based_rare_all<-bind_rows(ind_based_rare_all) # convert from list to data frame

colnames(ind_based_rare_all)<-c("Individuals","Richness","Year_season","site_id") # change columns names

# plot

ggplot(ind_based_rare_all,aes(x=Individuals,y=Richness,group = site_id ,color = site_id))+
  geom_line()

IndiP <- ggplot(ind_based_rare_all,aes(x=Individuals,y= Richness,color = site_id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
IndiP
```
###Group based individual rarefaction

```{r}


ind_based_rare<-list()

for (i in unique(group_year_data$site_id)) {

one_year<-group_year_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,7:ncol(one_year)]

rarefaction <- rarefaction.individual(year_sp_matrix, method = "sample-size", q = 0)

rarefaction$site_id<-i

ind_based_rare[[i]]<- rarefaction
  
}

ind_based_rare<-bind_rows(ind_based_rare)

colnames(ind_based_rare)<-c("Individuals","Richness","site_id")

ggplot(ind_based_rare,aes(x=Individuals,y=Richness,color = site_id))+
  geom_line(size = 1.5)+ scale_color_manual(values = color6) + theme_classic()
```

###Ordinations on log base 10 due to siganiid densities

```{r}
library(ggfortify)

#remove outlier Achziv 45 sample
Outlier_ac4516792 <- my_data %>%
  filter(date_id == "16792") %>%
   filter(depth_category >10)
my_data = my_data[!my_data$date_id == "16792",]
species_matrix <- my_data[,first_species:ncol(my_data)]
meta_data <- my_data[,1:29] # columns i want to keep from my data

log_trans_data <- decostand (species_matrix, method = "log",base = 10)
helge_trans_data <- decostand (species_matrix, method = "hellinger")
log10_trans_hel <- decostand (helge_trans_data, method = "log",base = 10)
```

my_data_complete   <- my_data[complete.cases(my_data), ]
colnames(my_data_complete)

my_data_complete <- my_data_complete[, colSums(my_data_complete != 0) > 0]
colnames(my_data_complete)
species_matrix_c <- my_data_complete[,first_species:ncol(my_data_complete)]
meta_data_c <- my_data_complete[,1:31] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_matrix_c, method = "hellinger")
log10_trans_hel_c <- decostand (helge_trans_data_c, method = "log",base = 10)
```{r}
rda_data <-  my_data%>%
  select(1:19,32:73)# columns i want to keep from my data
colnames(rda_data)
rda_data_complete   <- rda_data[complete.cases(rda_data), ]
rda_data_complete$site_id <-as.factor(rda_data_complete$site_id)  
species_rda_c <- rda_data_complete[,20:61]
meta_rda_c <- rda_data_complete[,1:19] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_rda_c, method = "hellinger")
log10_trans_hel_c <- decostand (helge_trans_data_c, method = "log",base = 10)


rda_results <- rda(log10_trans_hel_c ~ lat + depth_category+ Hard.Substrate +Soft.Substrate + Algae + reserve,
                data = rda_data_complete)

rda_results
```


```{r}
ordistep(rda_results ,na.action = na.omit)
```
```{r}
summary (rda_results)
```


Remember. When you see "Inertia", think "Variance". When you see "constrained" think "explained". So, in this example you can see that a proportion of 0.1121 of the inertia is explained. In words, the site_id,Site, knoll surface area and  depth affects ~10% of the all the community variance.


now lets plot the RDA using the `plot` function.
useful terminology for RDA plots:  

* sites = the scores (coordinates) or each row (i.e sample)    
* species = the unconstrained variables, actual species in our case    
* bp = bi plot, the effect of each constrained variable on the data    
* cn = centroid of non-continues constrained variable (for example Site or season)   
### RDA significance


The significance of your RDA can be tested using the function `anova.cca` which preform Anova like permutation tests. 

* don't be confused by the `cca` in the function name it can operate also on RDA

*Use either significance testing or model selection approch - not both!*
```{r}

ordiplot(rda_results, scaling = 2)
```

```{r}



colvec <- color6 #  colors according to your group



plot(rda_results,xlim = c(-2,2), 
     ylim = c(-3,3),
     # label the plot (title, and axes)
     main = "Biplot RDA - scaling 2",
     xlab = paste0("RDA1 (11.6) %"), 
     ylab = paste0("RDA2 (5.2%)") ) # plot RDA

with(rda_data_complete, points(rda_results,
                     display = "sites", col = colvec[site_id],
                     pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores

with(rda_data_complete, legend("bottomright", legend = levels(site_id), bty = "n",
                       col = colvec, pch = 21, pt.bg = colvec)) # add legend

#orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

text(rda_results, display = "cn", cex = 1, col = "blue") # add labels to centroieds


```


#The position of each species point in the ordination plot is determined by the strength and direction of the relationship between that species and the environmental variables included in the analysis. 
```{r}
plot(rda_results,xlim = c(-2,2), 
     ylim = c(-3,3),
     # label the plot (title, and axes)
     main = "Biplot RDA - scaling 2",
     xlab = paste0("RDA1 (11.6) %"), 
     ylab = paste0("RDA2 (5.2%)") ) # plot RDA

with(rda_data_complete, points(rda_results,
                     display = "sites", col = colvec[site_id],
                     pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores

with(rda_data_complete, legend("bottomright", legend = levels(site_id), bty = "n",
                       col = colvec, pch = 21, pt.bg = colvec)) # add legend

#orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

text(rda_results, display = "species", cex = 1, col = "blue") # add labels to centroieds


```
###Creating complete case data for temp diff and complexity measures
```{r}


my_data_complete   <- my_data[complete.cases(my_data), ]
colnames(my_data_complete)

my_data_complete <- my_data_complete[, colSums(my_data_complete != 0) > 0]
colnames(my_data_complete)
species_matrix_c <- my_data_complete[,first_species:ncol(my_data_complete)]
meta_data_c <- my_data_complete[,1:31] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_matrix_c, method = "hellinger")
log10_trans_hel_c <- decostand (helge_trans_data_c, method = "log",base = 10)
```
```{r}
rda_results_CC <- rda(log10_trans_hel_c ~ lat + mean_depth + TempDiff+ reserve +Rugosity +Curvature,
                data = my_data_complete)

rda_results_CC
```


```{r}
ordistep(rda_results_CC ,na.action = na.omit)
```
###Pairwize tukey

```{r}
nmds_data<-vegdist(log10_trans_hel,method = "bray")

pairwise<-betadisper(nmds_data, # the distance matrix
                     log10_hel_my_data$site_id) # the groups
# Reset graphics device state before opening a new PNG device
graphics.off()
# Open a PNG device with a larger size (e.g., 800x600 pixels)
par(oma = c(0, 80, 0, 0))
# Open a PNG device with a larger size
png("tukeyhsd_plot.png", width = 800, height = 600)

# Plot the TukeyHSD results with rotated y-axis labels
par(mar=c(3,10,2,1))
plot(TukeyHSD(pairwise), las = 1)



dev.off()


```


```{r}
ord  <- metaMDS(nmds_data,trace = FALSE)
```

  
We can see the Shepard plot of the nMDS. This described the "stress", Or, how well the nMDS is doing keeping the dissimilarity between points the same.  
 


*BTW - this plot is like assumption tests in models you should look at it but you don't need to present it in your final project...*

Large scatter around the line suggests that original dissimilarities are not well preserved in the reduced number of dimensions. You want to keep the stress as low as possible (anything above 0.2 is suboptimal).


Now let's check out the stress:

```{r}
ord$stress

```

OK, we preformed the analyses **now we can do 2 things:**

**1.** Visualize our results by plotting the nMDS  
**2.** Run statistic analyses

### 1. Visualize nMDS

Let's plot:

```{r }

plot(ord)

```

We can add some more information to the plot like different color for the different groups and convexhull 


* Note - you need ro run the lines together. Either mark all of them and than press `run` or `ctrl`+`enter` or (in r markdown) use the green arrow in the upper corner named -  `run current chunk`

```{r message=FALSE, warning=FALSE}

ordiplot(ord, type = "n",main = paste("stress=",round(ord$stress,3))) # create the plot
orditorp(ord, label = T, display = "sites" ,pch = 16) # add scores and color by sites
ordihull(ord, groups = my_data$site_id, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color6) # add convex hull
#legend("topleft", legend = levels(my_data$depth_category), bty = "n", col = colvec, pch = 15,cex=1.1) # add legend




the answer is Yes - Fish communities differ by site_id.  

if we want to know which survey were significantly different from one-another we can use post-hoc test and apply pair-wise comparison between the diffrent years.  

```{r}

pairwise<-betadisper(nmds_data, # the distance matrix
                     my_data$site_id) # the groups

par(mar=c(5,15,4,1))
plot(TukeyHSD(pairwise),las=2)

```

#### Simper

The simper functions performs pairwise comparisons of groups (i.e, `site_id`) and finds the contribution of each species to the average between-group Bray-Curtis dissimilarity.

**meaning** - which species contributed to the differences found between surveys.


```{r, results='hide'}
simper_output<-simper(log10_trans_hel, #my community data (transformed)
                      group = my_data$site_id) # my group

summary(simper_output)
```


```{r}
listAch25<- fish_filter%>%
  filter(site_id == "Achziv25") %>%
  select(15) %>%
  c(unique()
listAch25

listAch45<- fish_filter%>%
  filter(site_id == "Achziv45") %>%
  select(15) %>%
  unique()
listAch45
uniqueAch25 <-setdiff(listAch25,listAch45)
uniqueAch45 <-setdiff(listAch45,listAch25)


```

