---
title: "Data processing for chapter 3: 25-45 m sites" MKMRS data of fish survey  and invertebrate photoquadrates
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
WHEN UPDATING A NEW FISH_SURVEY FILE- CORRECT VIVIBILITY TO 7 AT AchzivR453Spring2020
AchzivR454Spring2020

---

### step 1 load data library etc and set wd


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
# library(assertthat)
# library(devtools)
# #For Diversity calculations
# library(plotrix)
# library(betapart)
library(here)
# library(ggridges)

# library(ggtext)# for italics in the plot legend 
# library(ggpubr) 
# Richness
library(rareNMtests)
library(mobr)

here<-here::here

rm(list = ls())
```

```{r}
# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")

# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
complex.inic <-read.csv(here("data", "EnvVar10_45.csv"))

fish_data <-merge(fish_data,sites, by= "site_id") 

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))
#Make sure Achziv is Reserve and Sdot Yam Not

head(fish_data)

color6 <- c("lightblue3", "blue2", "darkblue","yellow3", "darkorange", "brown")#  colors according to your group


```

###QA
```{r}

# Group by "tag" and count unique "sample_date" values
result <- fish_data %>%
  group_by(tag) %>%
  summarise(unique_dates = n_distinct(sample_date)) %>%
  filter(unique_dates > 1)

# Print the result
print(result)
```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs
Filtering cryptic species and species in the genus recognition level

```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter<-fish_data %>% filter(site=="Achziv"| site=="SdotYam" )
visibility_distribution <- table(fish_filter$visibility)
print(visibility_distribution)

fish_filter <-   fish_filter%>%
  filter(distance < 5.1)%>%
  filter(species != "Epinephelus sp") %>%
  filter(species != "Siganus sp") %>%
   filter(species != "Dasyatis.sp" ) %>%
  filter(species != "Empty transect") %>%
   filter(species != "Gobius sp" ) %>%
  filter(species != "Gobius bucchichi") %>%
  filter(species != "Gobius vittatus") %>%
   filter(species != "Dasyatis sp") 
fish_filter$reserve <- ifelse(fish_filter$site == "Achziv", "R", (ifelse(fish_filter$site_id == "SdotYam45", "N", "NR"))) # if its true write R if its false write Fall
head(fish_filter)


 # filter(season == "Fall")
```



##before we will convert the data to the wide format and start plotting we will add two more summary columns that will come to use later on. lets calculate the total number of individuals per survey and the total number of species.
#A. calculate the total number of individuals per species 
```{r}
fish_long <- fish_filter %>%
  dplyr::select(-transect,-a_b,-habitat, -depth, -factor_visibility, -observer, -size,-distance) %>%
  dplyr::relocate(species,.after = tag) %>%
   dplyr::mutate( occ =1 )

Abundance <- fish_long %>%
  filter(site == "Achziv" | site == "SdotYam")%>%
    dplyr::group_by(across(site_id:date_id)) %>% 
  dplyr::summarise(Abundance = sum(occ)) 
head( Abundance)
fish_long <- fish_long %>% 
  dplyr::group_by(across(site_id:species)) %>% 
 dplyr::summarise(total_ind = sum(occ))
```
#plotting abundance 
```{r}


Plot_abundance <- Abundance%>%
  ggplot(aes(y = Abundance)) +
  geom_boxplot(notch=TRUE, notchwidth = 0.8) +
  facet_wrap(~site_id) +
  theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Abundance per 1000 square m") + theme(axis.text.x = element_blank())
Plot_abundance_hist <- Abundance%>%
  ggplot(aes(y = Abundance)) +
  geom_histogram(notch=TRUE, notchwidth = 0.8) +
  facet_wrap(~depth_category) +
  theme_classic() +
  scale_color_manual(values = color6) +
  ylab("Abundance per 1000 square m") + theme(axis.text.x = element_blank())
Plot_abundance_hist


print(Plot_abundance)


```
#testing abundance
```{r}

your_data <- Abundance   %>%
  dplyr::filter(site == "Achziv" ) %>% 
     dplyr::filter(depth_category>10)
kruskal.test(Abundance ~ site_id, data = your_data)
kruskal.test(visibility ~ depth_category, data = your_data)
your_data <- Abundance   %>%
  dplyr::filter(site == "SdotYam" ) %>% 
     dplyr::filter(depth_category>10)
kruskal.test(Abundance ~ site_id, data = your_data)
kruskal.test(visibility ~ depth_category, data = your_data)
```



#B. Calculate  the total number of species (n_species_sample)and total individuals per survey (n_ind_sample))

```{r}
fish_long<-fish_long %>%
  dplyr::group_by(tag)%>% 
  dplyr::mutate(n_species_sample = n_distinct(species),
          n_ind_sample= sum(total_ind)) %>%
  dplyr::ungroup()
head(fish_long)
```

###Finally, lets convert our data to wide format:
```{r}
fish_wide<- spread(fish_long,species,total_ind,fill = 0)
head(fish_wide)
colnames(fish_wide)


```
## Is visibility different in Sdot Yam vs Achziv? visibility testing
```{r}
library(car)



# Visual check 
ggplot(fish_long, aes(x=visibility)) +
  geom_histogram(binwidth=1, fill="skyblue", color="black") +
  labs(title="Histogram of Visibility",
       x="Visibility",
       y="Frequency")+
  facet_grid(site ~ depth_category) +
  ggtitle("Frequency Plots for Visibility by Site and Depth Category")
#Shapiro-Wilks testing
df <-fish_wide %>%
  group_by(site, depth_category) %>%
  summarize(shapiro_p = shapiro.test(visibility)$p.value)
df # NOt normally distributed
```
#A-parametric testing visibility
```{r}
library(rstatix)
df <-fish_wide%>%
  filter(site %in% c("Achziv", "SdotYam")) %>%
  group_by(depth_category) %>%
  wilcox_test(visibility ~ site) %>%
  adjust_pvalue(method = "bonferroni")
df
```

```{r}
ggplot(fish_long, aes(sample = visibility)) +
  geom_qq() +
  geom_qq_line() +
  facet_grid(site ~ depth_category) 
#with Q-Q plots
ggplot(fish_long, aes(sample = visibility)) +
  geom_qq() +
  geom_qq_line() +
  facet_grid(site ~ depth_category) +
  ggtitle("Q-Q Plots for Visibility by Site and Depth Category")
```
###making Fish abudance table filtering empty species

```{r}


fish_wide <- fish_wide[, colSums(fish_wide != 0) > 0]

write.csv(fish_wide, here("R.generated.Data", "45Fish_data_clean.csv"))

```

### Invertebrate summary according to taxonomy_group : allFish10.45dataFact


```{r}

Inv.Grp.data <- read.csv(here( "Data","inv_group_mat_wis.csv"))
Inv.Grp.data = subset(Inv.Grp.data, select = -c(tag) )
Inv.Grp.data<- Inv.Grp.data %>% mutate (tag= paste0(site
,depth,transect,season,year))
colnames(Inv.Grp.data)
Inv.Grp.factors <- dplyr::select(Inv.Grp.data,Algae,Bryozoan,Hard.Substrate, Hydrozoa, Soft.Substrate,Sponge,tag, 11:19)

head(Inv.Grp.factors)

DepthFishdataINv <- dplyr::select(Inv.Grp.factors, 1:7)


```
###Combine Rock cover wiskonsin transformed and complexity factors into the DepthFishdataFact df
```{r}

DepthFishdataFact <-left_join(fish_wide, DepthFishdataINv, by = "tag")
DepthFishdataFact <-left_join(DepthFishdataFact, complex.inic, by = "tag")

colnames(DepthFishdataFact)
DepthFishdataFact <-DepthFishdataFact %>%
  relocate("Algae" :"TempDiff",.after = tag) 
head(DepthFishdataFact)
colnames(DepthFishdataFact)


```
###Resolving date discrepencies between Fish and epibenthic cover data sets
```{r}



checktag<-dplyr::select(DepthFishdataFact,4,20,13)
#view(checktag)
# Update sample_date.x column where sample_date.y is not NA and is different
DepthFishdataFact$sample_date.x[!is.na(DepthFishdataFact$sample_date.y) & DepthFishdataFact$sample_date.y != DepthFishdataFact$sample_date.x] <- DepthFishdataFact$sample_date.y[!is.na(DepthFishdataFact$sample_date.y) & DepthFishdataFact$sample_date.y != DepthFishdataFact$sample_date.x]
#remove extra date column
DepthFishdataFact <- DepthFishdataFact[, !names(DepthFishdataFact) %in% "sample_date.y"]

# Print the updated dataframe to verify the changes
colnames(DepthFishdataFact)
DepthFishdataFact$sample_date <- DepthFishdataFact$sample_date.x
DepthFishdataFact$sample_date.x <- NULL
DepthFishdataFact <-DepthFishdataFact %>%
  relocate(sample_date,.after = sample_year) 
write.csv(DepthFishdataFact, here("R.generated.Data","my_data.csv"))

```

```{r}
colnames(DepthFishdataFact)

corrDF <-subset(DepthFishdataFact,select =c(8,14:19,25:26,28,29 ))
cor_matrix <- cor(corrDF, use = "complete.obs", method = "spearman")
# Print the correlation matrix
corrplot::corrplot(cor_matrix, 
         type = "upper",
         tl.col = "black",  # Text label color
         tl.srt = 45,,       # Rotate text labels
         diag = FALSE)      # Rotate text labels)
corrplot::corrplot(cor_matrix, 
         method = "color",  # Use colored squares
         type = "upper",    # Show only upper triangle
         order = "hclust",  # Order variables by hierarchical clustering
         addCoef.col = "black",  # Add correlation coefficients
         tl.col = "black",  # Text label color
         tl.srt = 45,       # Rotate text labels
         diag = FALSE,      # Don't show correlations on the diagonal
         # Custom color palette for negative to positive correlations
         col = colorRampPalette(c("#6D9EC1", "white", "#E46726"))(200),
         # Adjust this to change the intensity of the colors
         cl.lim = c(-1, 1))

```

####Diversity
```{r}

color6 = c("lightblue3", "blue2", "darkblue")
my_data <- DepthFishdataFact %>%
  mutate(Year_season_id =paste0(sample_year,season,site_id)) %>%
  relocate("Year_season_id",.after = tag) 
 head(my_data)
 colnames(my_data)
```


###Species Matrix +Metadata +transformations
```{r}
colnames(my_data)
first_species <- 32
sp_matrix <- my_data[,first_species:ncol(my_data)] %>%
  select_if(~ sum(.) != 0)
meta_data <- my_data[,1:31] # columns i want to keep from my data


hel_trans <- decostand (sp_matrix, method = "hellinger" )
log2_trans_hel <- hel_trans#decostand (hel_trans,method = "log" ,base = 2)
hel_my_data <- bind_cols(my_data[, -( first_species: ncol(my_data))], log2_trans_hel)
colnames (hel_my_data)
```
### Hill numbers profile Allways use  hill = T
```{r}
renyi_profile <- renyi(sp_matrix,  
                       scales = c(0,  1.0, 2),
                       hill = T)
head(renyi_profile)
metadataRug <- select(meta_data,1:14, 21:28)
renyi_df<-bind_cols(metadataRug,renyi_profile)

renyi_df<-gather(renyi_df,"Q","Value",23:25)
head(renyi_df)
renyi_df<-renyi_df %>% arrange(site_id,depth_category,sample_year, season,tag) %>%
   filter(!is.na(Q))

renyi_df$Q <- as.numeric(renyi_df$Q)
```
#Diversity with rugosity VR complete case Mixed effects modelling
```{r}
cmplex_data_complete   <- renyi_df[complete.cases(renyi_df), ]
head (cmplex_data_complete)
```

###Testing Species diveristy is affected by , lat
Start by loading the data and having a look at them.

```{r}
Q1df <-cmplex_data_complete %>% 
  filter(Q==1)
head(Q1df)
hist(Q1df$LogVR)  # seems close to normal distribution - good!

#decostand (Q1df$Value, method = "sq")
hist(Q1df$Value)  # seems far from normal distribution -transform

Q1df <- Q1df %>% 
 mutate(SqValue = (Value)^0.5)
hist(Q1df$SqValue) # seems close to normal distribution - good!
Q1df$lat <- as.factor(Q1df$lat)
```

#Centering the data
```{r}
#we first center and scale the data. this is often useful for mixed-effect models. note that we are only using this for the stats, not to show them in the graphs

Q1df$LogVRCent <- scale(Q1df$LogVR, center = TRUE, scale = TRUE)
Q1df$mean_depth <- scale(Q1df$LogVR, center = TRUE, scale = TRUE)
basic.lm <- lm(Value~ LogVRCent, data = Q1df)

summary(basic.lm)

## Let's plot the data with ggplot2

ggplot(Q1df, aes(x = LogVRCent, y = mean_depth)) +
  geom_point()+
  geom_smooth(method = "lm")
```
```{r}
if (!require(pacman)) install.packages('pacman')
library(pacman)
p_load(ggplot2, lme4,lmerTest,MuMIn)

Q1df$sample_year <- as.factor(Q1df$sample_year)

mixed.lmer <- lmer(Value ~ LogVRCent * mean_depth +  (1|sample_year), data = Q1df) # this is how you write random component effects only the intercept : after this | symbol; assigning slope to the random variables means that you suspect that there is an interaction between the factor and the affected variable
summary(mixed.lmer)
r.squaredGLMM(mixed.lmer) #Marginal r-square (R2m) gives the proportion of the fixed in relation to all variation; the conditional r square(R2c) (conditional on the specific beaches we included) includes all the explained variation (fixed an d random). We report the marginal R square

#Random effect
VarCorr(mixed.lmer)

# Fixed effect
fixef(mixed.lmer) #compare these to the estimates of the simple lm model!
#summary(basic.lm)
# b0 & b1 for Individual sites 
coef(mixed.lmer)

```

###SEction 2: Grouping per sampling day
#Grouping long format
```{r}
colnames
head(fish_long)
Lgroup_year_data <- fish_long %>% 
  mutate(Year_season_id =paste0(sample_year,season,site_id)) %>%
  relocate("Year_season_id",.after = tag)   %>%
  dplyr::select(site,date_id, Year_season_id, site_id, depth_category, season, species, total_ind) %>% 
  dplyr::group_by(site,date_id, Year_season_id, site_id, depth_category, season, species) 
  Lgroup_year_data <- Lgroup_year_data %>% 
   dplyr::group_by(site,date_id, Year_season_id, site_id, depth_category, season, species )  %>%  mutate(occ = n_distinct(species, na.rm = TRUE)) %>%  dplyr::group_by(Year_season_id) %>% mutate(no.species = sum(occ, na.rm = TRUE), total_gr_sp = sum(total_ind))
head(Lgroup_year_data)
colnames(Lgroup_year_data)
write.csv(Lgroup_year_data, here("R.generated.Data","grouped_long.csv"))
```



#Grouping wide format 
```{r}
colnames(my_data)
group_year_data <- my_data %>% 
  dplyr::select(date_id, Year_season_id, site_id, depth_category, season, "Acantholabrus palloni" :"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni" :"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_year_data)
save(group_year_data, file = "group_year_data.RData")
colnames(group_year_data)

colnames(hel_my_data)
group_trans_data <- hel_my_data %>% 
  dplyr::select(date_id, sample_year,site, Year_season_id, site_id, depth_category, season, "Acantholabrus palloni" :"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id, site, sample_year, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni" :"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_trans_data)



```
###Gamma, beta, alfa

```{r}

group_year_matrix <- group_year_data %>%
  ungroup()  %>% 
  dplyr::select("Acantholabrus palloni" :"Torquigener flavimaculosus") 

gamma_diversity <- renyi(group_year_matrix, scales = c(0,  1.0, 2),
                       hill = T)
colnames(group_year_data)
metadataX <- select(group_year_data,2,3)
renyi_df2<-bind_cols(metadataX,gamma_diversity)
colnames(renyi_df2)
renyi_df3<-gather(renyi_df2,"Q","Value",6:8)

```

#Plotting gamma diversity hill numbers
```{r}
QPlot <- renyi_df3 %>% 
  ggplot(aes(y = Value, x = as.factor(depth_category), fill = as.factor(depth_category))) +
  geom_boxplot(notch = TRUE, notchwidth = 0.8) +
  labs(y = "Hill numbers", x = "Depth (m)") +
  theme_classic() +
  facet_wrap(~Q) +
  scale_fill_manual(values = color6) + theme_minimal()+
  theme(legend.position = "none")  # Remove legend if not needed

QPlot
```

#testing gamma diversity




```{r}

df <-renyi_df3%>%
  group_by(Q) %>%
  wilcox_test(Value ~ depth_category) %>%
  adjust_pvalue(method = "bonferroni")
df
```



```{r}
site_id_distribution <- table(group_year_data$site_id,group_year_data$season )
print(site_id_distribution)
```


```{r}
least_sampled <- 23

least_sampled
```



##Incidence data- beta pairs
```{r}
AC_incidence <- sp_matrix %>% 
  replace(sp_matrix > 0, 1) 
AC_incidence_bind<-bind_cols(meta_data,AC_incidence )


caves_beta_pairs_tr <- beta.pair(AC_incidence)
```


##Group Incidence data- beta pairs depth catego
```{r}
head(group_year_data)
sp_data_group <- group_year_data %>%
  ungroup()  %>% 
  select(6:48)
meta_group<-group_year_data %>%
  ungroup()  %>% 
  select(3)
AC_incidenceGrp <- sp_data_group %>%
  replace(sp_data_group > 0, 1) 
AC_incidence<-bind_cols(meta_group,AC_incidenceGrp )


caves_beta_pairs_gr <- beta.pair(AC_incidenceGrp)
```
#group beta pairs
```{r message=FALSE, results='hide'}


caves_beta_sample_gr <- beta.sample(AC_incidenceGrp, 
            sites= 5,  # how many knolls to randomly sample
            samples = 100) # how many repetitions of (10) knolls to do 


```


after we understood how the function is working lets calculate the multi-site beta diversity of all the years for the site_id   
###Sorensen beta diversity: per transect 

```{r message=FALSE, warning=FALSE, results='hide'}


caves <-my_data  

caves[first_species:ncol(caves)] <-replace(caves[first_species:ncol(caves)],
                                            caves[first_species:ncol(caves)] > 0,
                                            1) # incident data

caves_all_years<-list() # empty list

for (i in unique(caves$depth_category)) { # loop
  
  one_year <-  caves %>% filter(depth_category == i)
  
  one_year_sp_matrix <- one_year[,first_species:ncol(one_year)]
  
  caves_beta_sample <- beta.sample(one_year_sp_matrix,
            sites= 5, 
            samples = 100)
  
 
  caves_beta_sample_data <- data.frame(mean = caves_beta_sample[["mean.values"]],
                                    se = caves_beta_sample[["sd.values"]])

  caves_beta_sample_data <-rownames_to_column(caves_beta_sample_data,'Component') 

  caves_beta_sample_data$depth_category <- i
  
 caves_all_years[[i]]<-caves_beta_sample_data
  
}


caves_all_years<-bind_rows(caves_all_years)


```


```{r}
caves_all_years<-caves_all_years %>% filter(Component != "beta.SOR")

TransectSOr<- ggplot(caves_all_years)+
  aes(x=depth_category,y=mean,fill=Component)+
  geom_bar(stat="identity",position = position_stack())+
  ylab(expression(paste(beta,"-diversity" (Sorensen)))) +theme_classic() + scale_y_continuous(limits = c(0, 0.8))

TransectSOr


```
###Sorensen beta diversity:  grouped

```{r message=FALSE, warning=FALSE, results='hide'}


caves <-group_year_data  
first_species = 6
caves[first_species:ncol(caves)] <-replace(caves[first_species:ncol(caves)],
                                            caves[first_species:ncol(caves)] > 0,
                                            1) # incident data

caves_all_years<-list() # empty list

for (i in unique(caves$depth_category)) { # loop
  
  one_year <-  caves %>% filter(depth_category == i)
  
  one_year_sp_matrix <- one_year[,first_species:ncol(one_year)]
  
  caves_beta_sample <- beta.sample(one_year_sp_matrix,
            sites= 5, 
            samples = 100)
  
 
  caves_beta_sample_data <- data.frame(mean = caves_beta_sample[["mean.values"]],
                                    se = caves_beta_sample[["sd.values"]])

  caves_beta_sample_data <-rownames_to_column(caves_beta_sample_data,'Component') 

  caves_beta_sample_data$depth_category <- i
  
 caves_all_years[[i]]<-caves_beta_sample_data
  
}


caves_all_years<-bind_rows(caves_all_years)


```


```{r}
caves_all_years<-caves_all_years %>% filter(Component != "beta.SOR")

PGrTransSor<- ggplot(caves_all_years)+
  aes(x=as.factor(depth_category),y=mean,fill=Component)+
  geom_bar(stat="identity",position = position_stack())+
  ylab(expression(paste(beta,"-diversity " (Sorensen)))) +theme_classic() + scale_y_continuous(limits = c(0, 0.8))+xlab("Depth category")

PGrTransSor
```


```{r}
library(ggpubr)
# Set a common x-axis titlec ommon.legend = TRUE,  common.xlabel = "N" )
tiff("Betaplots.jpeg", units="in", width=10, height=6, res=600)
betaPlots <- ggarrange( TransectSOr, PGrTransSor, 
          labels = c("    Transect","   Sample day"),
          ncol = 2, nrow = 1, common.legend = TRUE)   # Set a common x-axis title)
betaPlots
dev.off()
betaPlots
```

We will use the function `rarefaction.sample` from the package `rareNMtests`. 

In this function the first argument is the data name,  the method argument `sample-size` means that the function will use incidence data and `q = 0` is used for species richness rarefaction.  

##Grouped sample based rarefaction Richness

```{r}
sample_based_rare<-list()

for (i in unique(group_year_data$depth_category)) {

one_year<-group_year_data %>% filter(depth_category == i)
year_sp_matrix <- one_year[,6:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q =0)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Richness","Depth_category")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

Rp <- ggplot(sample_based_rare,aes(x=Samples,y= Richness,color = as.factor(Depth_category)))+geom_line(size = 1.2)  +theme_classic() + scale_color_manual(values = color6)  +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)# Enlarge legend text
  ) 
Rp
```

```{r}
# Install and load required packages

library(iNEXT)
library(ggplot2)

# Assuming you have abundance data in a vector called 'abundance_data'
# If not, you can use the example data from iNEXT

sp_data_groupIN<- sp_data_group %>% 
  ungroup()  %>%mutate ("sampling_unit" = 1)%>% relocate(sampling_unit)%>% as.data.frame()
# Run iNEXT
out <- iNEXT(sp_data_groupIN, q = 0, datatype = "incidence_freq", endpoint = NULL)

# Generate the rarefaction plot with confidence intervals
ggiNEXT(out, type = 1, se = TRUE) +
  theme_bw() +
  labs(x = "Number of individuals", y = "Species richness") +
  theme(legend.position = "none")
```





#Grouped sample base Effective diversity
```{r}
sample_based_rare<-list()

for (i in unique(group_year_data$depth_category)) {

one_year<-group_year_data %>% filter(depth_category == i)
year_sp_matrix <- one_year[,6:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q =1)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Effective.diversity","Depth_category")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

EffDp <- ggplot(sample_based_rare,aes(x=Samples,y= Effective.diversity,color = as.factor(Depth_category)))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) +ylab("Effective diversity")
EffDp

```
#Grouped sample base Real diversity
```{r}
sample_based_rare<-list()

for (i in unique(group_year_data$depth_category)) {

one_year<-group_year_data %>% filter(depth_category == i)
year_sp_matrix <- one_year[,6:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q =2)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","True.diversity","depth_category")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

RealDp <- ggplot(sample_based_rare,aes(x=Samples,y= True.diversity,color = as.factor(depth_category)))+geom_line(size = 1.2) + scale_color_manual(values = color6) +ylab("True diversity") +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
RealDp

```


```{r}
RarePlots <- ggarrange( Rp, EffDp,RealDp,
                    ncol = 3, nrow = 1, common.legend = TRUE)  # Set a common x-axis title) #labels = c("Richness ", "Effective diveristy", "True diversity")

 RarePlots

```

```{r warning=FALSE,results='hide'}
colnames(group_year_data)
Grpupedsp_matrix<-group_year_data %>%
  ungroup() %>% 
  filter(depth_category== 45| depth_category== 25)%>%
  select(site_id,  "Acantholabrus palloni"  :ncol(group_year_data)) %>% arrange(site_id)

Grpupedsp_matrix<-as.data.frame(Grpupedsp_matrix) # make sure your data is defined as data.frame

eco_test <- EcoTest.sample(Grpupedsp_matrix[,-1], by = Grpupedsp_matrix$site_id, niter = 500, MARGIN=1,q=2, trace = FALSE)


```


```{r}
plot.new()
plot(eco_test)

```




```{r}
# check<-real_group %>% mutate(total_abundance = rowSums(real_group[2:length(real_group)]),
#          total_rihcness = rowSums(real_group[2:length(real_group)]>0)) %>% 
#   select(site_id,total_abundance,total_rihcness)
#   

```

```{r}
library(ggfortify)
rm(list = setdiff(ls(), c("my_data", "group_trans_data")))

#remove outlier Achziv 45 sample
Outlier_ac4516792 <- my_data %>%
  filter(date_id == "16792") %>%
   filter(depth_category >10)
my_data = my_data[!my_data$date_id == "16792",]

species_matrix <- my_data[,32:ncol(my_data)]
meta_data <- my_data[,1:31] # columns i want to keep from my data



log_trans_data <- decostand (species_matrix, method = "log",base = 2)
helge_trans_data <- decostand (species_matrix, method = "hellinger")
log2_trans_hel <- helge_trans_data#decostand (helge_trans_data, method = "log",base = 2)
```
#Group data prep
```{r}
colnames(group_trans_data)
first_species=8
group_trans_data = group_trans_data[!group_trans_data$date_id == "16792",]
species_matrix <- group_trans_data[,first_species:ncol(group_trans_data)]
meta_data <- group_trans_data[,1:7] # columns i want to keep from my data

log_trans_data <- decostand (species_matrix, method = "log",base = 2)
helge_trans_data <- decostand (species_matrix, method = "hellinger")
log2_trans_hel <-helge_trans_data #decostand (helge_trans_data, method = "log",base = 2)

```



###Dat preparation for RDA models A. Group complete epibentich +fish per transect
```{r}
colnames(my_data)
group_year_data <- my_data %>% 
  dplyr::select(date_id, site, Year_season_id, site_id, depth_category,sample_year, season, "Acantholabrus palloni" :"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id,site, site_id,depth_category,sample_year, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni" :"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_year_data)
colnames(group_year_data)
rda_data <-  group_year_data

rda_data_complete   <- group_year_data[complete.cases(group_year_data), ]
rda_data_complete$site_id <-as.factor(rda_data_complete$site_id)
colnames(rda_data_complete)
species_rda_c <- rda_data_complete[,8:ncol(rda_data_complete)]
meta_rda_c <- rda_data_complete[,1:7] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_rda_c, method = "hellinger")
hel_trans_data <- helge_trans_data_c#decostand (helge_trans_data_c, method = "log",base = 2)

colnames(rda_data_complete)
```
```{r}
rda_results <- rda(hel_trans_data~ site +season +  depth_category+sample_year,
                data = rda_data_complete)

rda_results



```

                
#CCA was tested but it explains even less of the variation.
cca_results <- cca(log2_trans_hel~ season + depth_category+ reserve+ Hard.Substrate +Soft.Substrate + Algae ,
                data = my_data,
                na.action = na.omit) 

plot(cca_results, scaling = 3)

cca_results


```{r}
sel.os <-ordistep(rda_results ,na.action = na.omit)
sel.os 
```



Remember. When you see "Inertia", think "Variance". When you see "constrained" think "explained". So, in this example you can see that a proportion of 0.1121 of the inertia is explained. In words, the site_id,Site, knoll surface area and  depth affects ~10% of the all the community variance.


now lets plot the RDA using the `plot` function.
useful terminology for RDA plots:  

* sites = the scores (coordinates) or each row (i.e sample)    
* species = the unconstrained variables, actual species in our case    
* bp = bi plot, the effect of each constrained variable on the data    
* cn = centroid of non-continues constrained variable (for example Site or season)   
### RDA significance


The significance of your RDA can be tested using the function `anova.cca` which preform Anova like permutation tests. 

* don't be confused by the `cca` in the function name it can operate also on RDA

*Use either significance testing or model selection approch - not both!*
```{r}
#hellinger
ordiplot(rda_results, scaling = 2)
orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species


```
```{r}
#Hellinger and Log transformed

ordiplot(rda_resultsLog, scaling = 2)
orditorp(rda_resultsLog,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

```


```{r}



colvec <- c("lightblue3", "blue2", "darkblue","yellow3", "darkorange", "brown")#  colors according to your group



plot(rda_resultsLog,xlim = c(-2,2), 
     ylim = c(-1,3),
     # label the plot (title, and axes)
     main = "Biplot RDA ",
     xlab = paste0("RDA1 (7.8) %"), 
     ylab = paste0("RDA2 (3.7%)") ) # plot RDA

with(rda_data_complete, points(rda_resultsLog,
                     display = "sites", col = colvec[site_id],
                     pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores

with(rda_data_complete, legend("bottomright", legend = levels(site_id), bty = "n",
                       col = colvec, pch = 21, pt.bg = colvec)) # add legend

#orditorp(rda_resultsLog,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

text(rda_resultsLog, display = "cn", cex = 1, col = "blue") # add labels to centroieds


```


#The position of each species point in the ordination plot is determined by the strength and direction of the relationship between that species and the environmental variables included in the analysis. 
```{r}
# plot(rda_results,xlim = c(-2,2), 
#      ylim = c(-3,3),
#      # label the plot (title, and axes)
#      main = "Biplot RDA - scaling 2",
#      xlab = paste0("RDA1 (10.8) %"), 
#      ylab = paste0("RDA2 (4.2%)") ) # plot RDA
# 
# with(rda_data_complete, points(rda_results,
#                      display = "sites", col = colvec[site_id],
#                      pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores
# 
# with(rda_data_complete, legend("bottomright", legend = levels(site_id), bty = "n",
#                        col = colvec, pch = 21, pt.bg = colvec)) # add legend
# 
# #orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species
# 
# text(rda_results, display = "species", cex = 1, col = "blue") # add labels to centroieds


```
###Creating complete case data for temp diff and complexity measures
```{r}


my_data_complete   <- my_data[complete.cases(my_data), ]
colnames(my_data_complete)

my_data_complete <- my_data_complete[, colSums(my_data_complete != 0) > 0]
colnames(my_data_complete)
species_matrix_c <- my_data_complete[,32:ncol(my_data_complete)]
meta_data_c <- my_data_complete[,1:31] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_matrix_c, method = "hellinger")
trans_hel_c <-helge_trans_data_c #decostand (helge_trans_data_c, method = "log",base = 2)
```
```{r}

rda_results_CC <- rda(trans_hel_c  ~  mean_depth + TempDiff + site + season + Rugosity + LogVR,
                data = meta_data_c)

rda_results_CC

```

```{r}
sel.osCC <-ordistep(rda_results_CC ,na.action = na.omit)

sel.osCC
```


###Pairwize tukey

```{r}
nmds_data<-vegdist(log2_trans_hel,method = "bray")

pairwise<-betadisper(nmds_data, # the distance matrix
                     group_year_data$site_id) # the groups
# Reset graphics device state before opening a new PNG device
graphics.off()

# Open a PNG device with a larger size
png("tukeyhsd_plot.png", width = 1000, height = 600, res = 100)

# Set margins and plot region
par(mar = c(5, 15, 4, 2) + 4)  # Increase left margin (second value)
par(plt = c(0.2, 0.95, 0.1, 0.9))  # Adjust plot region

# Plot the Tukey HSD results
plot(TukeyHSD(pairwise), las = 1, xpd = TRUE)

# Close the device
dev.off()








```


```{r}
ord  <- metaMDS(nmds_data,trace = FALSE)
```

  
We can see the Shepard plot of the nMDS. This described the "stress", Or, how well the nMDS is doing keeping the dissimilarity between points the same.  
 


*BTW - this plot is like assumption tests in models you should look at it but you don't need to present it in your final project...*

Large scatter around the line suggests that original dissimilarities are not well preserved in the reduced number of dimensions. You want to keep the stress as low as possible (anything above 0.2 is suboptimal).


Now let's check out the stress:

```{r}
ord$stress

```

OK, we preformed the analyses **now we can do 2 things:**

**1.** Visualize our results by plotting the nMDS  
**2.** Run statistic analyses

### 1. Visualize nMDS

Let's plot:

```{r }

plot(ord)

```

We can add some more information to the plot like different color for the different groups and convexhull 


* Note - you need ro run the lines together. Either mark all of them and than press `run` or `ctrl`+`enter` or (in r markdown) use the green arrow in the upper corner named -  `run current chunk`


#### Adonis

Assessing differences in community composition is done with *per*mutational *M*ultivariate *A*nalysis of *V*ariance, or *perMANOVA*. These tests are done on distances, meaning that they assess the differences between communities based on dissimilarity. With perMANOVA, the null hypothesis is that the centroids of your groups (in ordination space as defined by the dissimilarity measure you’ve chosen) are equivalent for all groups. In other words, you are asking if, following some measure of (dis)similarity, the community composition of sites between groups is the same.

This explanation is taken from the following website and contain a lot of useful information! 
https://rpubs.com/an-bui/vegan-cheat-sheet

After we decided what environmental variables we want to use we can test their contribution using `adonis2` function. 


if I want to know if fish communities significantly changed along the surveys I can ask it that way:

```{r}
adon.results <- adonis2(nmds_data ~ site_id, 
                        data = group_year_data,
                        method="bray",
                        perm=999,
                        na.action =na.omit)

adon.results


```

the answer is Yes - Fish communities differ by site id  

if we want to know which sites were significantly different from one-another we can use post-hoc test and apply pair-wise comparison between the sites.  



```{r}

pairwise<-betadisper(nmds_data, # the distance matrix
                     group_year_data$site_id) # the groups

par(mar=c(5,15,4,1))
plot(TukeyHSD(pairwise),las=2)

```

#### Simper

The simper functions performs pairwise comparisons of groups (i.e, `site_id`) and finds the contribution of each species to the average between-group Bray-Curtis dissimilarity.

**meaning** - which species contributed to the differences found between surveys.


```{r, results='hide'}
simper_data<-group_trans_data 
colnames(simper_data)
 species_matrix_simper <- simper_data[,8:ncol(simper_data)]

  simper_output <-simper(species_matrix_simper, #my community data (transformed)
                      group = simper_data$site_id) # my group


  summary (simper_output)


```

#Unique species by site
```{r}
site_id_fish <- fish_long %>% 
  filter(site =="Achziv" |site == "SdotYam")   
Td_fish<-  as.data.frame(table(site_id_fish$site_id, site_id_fish$species))%>%
  dplyr::filter(Freq>0)
 
 SdotYam45<- dplyr::filter(Td_fish,Var1 =="SdotYam45")

 SY45List <- as.character(unique(SdotYam45$Var2))
print(SY45List)

 SdotYam25<- dplyr::filter(Td_fish,Var1 =="SdotYam25")

 SY25List <- as.character(unique(SdotYam25$Var2))
print(SY25List)

```
###Trait data
```{r}
library(rfishbase)
library(dplyr)  # For data manipulation

# Your species list
SY45List <- c("Acantholabrus palloni", "Chromis chromis", "Coris julis", "Diplodus vulgaris", 
              "Epinephelus costae", "Epinephelus marginatus", "Muraena helena", "Mycteroperca rubra", 
              "Parupeneus forsskali", "Pterois miles", "Sparisoma cretense", "Sardin sp", 
              "Scomber sp", "Serranus cabrilla", "Siganus rivulatus", "Sparisoma cretense", 
              "Symphodus mediterraneus", "Symphodus tinca", "Synodus saurus", "Thalassoma pavo", 
              "Torquigener flavimaculosus")

# Function to safely get species info
get_species_info <- function(species_name) {
  tryCatch({
    info <- species(species_name)
    if (nrow(info) > 0) {
      return(info)
    } else {
      return(NULL)
    }
  }, error = function(e) {
    message("Error processing ", species_name, ": ", e$message)
    return(NULL)
  })
}

# Get info for SY45  species
species_info_list <- lapply(SY45List, get_species_info)

# Combine all data frames into one
SY45species_df <- do.call(rbind, species_info_list)
not.in.dist.df<- setdiff(SY45List,SY45species_df$Species) 
# Clean up the dataframe
SY45species_df <- SY45species_df %>%
  select(1, 28:33) %>%
  distinct()
SY45List <- c(
  "Atherinomorus forskalii", "Balistes capriscus",
  "Chromis chromis", "Coris julis",
  "Diplodus vulgaris", "Epinephelus costae",
  "Epinephelus marginatus", "Muraena helena",
  "Mycteroperca rubra", "Parablennius rouxi",
  "Parupeneus forsskali", "Pterois miles",
  "Sparisoma cretense", "Sardin sp",
  "Sargocentron rubrum", "Scomberomorus commerson",
  "Seriola dumerili", "Serranus cabrilla",
  "Serranus scriba", "Siganus luridus",
  "Siganus rivulatus", "Sparisoma cretense",
  "Symphodus mediterraneus", "Symphodus tinca",
  "Thalassoma pavo", "Torquigener flavimaculosus"
)
# Get info for SY25  species
species_info_list <- lapply(SY25List, get_species_info)

# Combine all data frames into one
SY25species_df <- do.call(rbind, species_info_list)
not.in.dist.df<- setdiff(SY25List,SY25species_df$Species) 
print(not.in.dist.df)
# Clean up the dataframe
SY25species_df <- SY25species_df %>%
  select(1, 28:33) %>%
  distinct()
# View the result
print(SY25species_df)

in.45.in.25<- setdiff(SY45List,SY25species_df$Species) 
print(in.45.in.25)
SYspecies_in.45not.in.25 <- lapply(in.45.in.25, get_species_info)
print(SYspecies_in.45not.in.25)
SIMPER_Ach45 <-c("Sciaena umbra","Diplodus sargus", "Scarus ghobban")
SIMPER_Ach45_info_list <- lapply(SIMPER_Ach45, get_species_info)
SIMPER_Ach45_info_list <- do.call(rbind, SIMPER_Ach45_info_list)
SIMPER_Ach45_info_list <- SIMPER_Ach45_info_list %>%
  select(1, 28:33) %>%
  distinct()
# View the result
print(SIMPER_Ach45_info_list)
```


```{r}
# Get info for SY45  species
species_info_list <- lapply(SY45List, get_species_info)

# Combine all data frames into one
SY45species_df <- do.call(rbind, species_info_list)
not.in.dist.df<- setdiff(SY45List,SY45species_df$Species) 
# Clean up the dataframe
SY45species_df <- SY45species_df %>%
  select(1, 28:33) %>%
  distinct()

# View the result
print(SY45species_df)


uniqueSY45 <-setdiff(SY45List,SY25List)
uniqueSY45
uniqueAch45 <-setdiff(listAch45,listAch25)


```



