---
title: "Data processing for chapter 3: 25-45 m sites" MKMRS data of fish survey  and invertebrate photoquadrates
output: html_document
date: "2022-11-10"
preparing 3 data sets and combined data sets:
WHEN UPDATING A NEW FISH_SURVEY FILE- CORRECT VIVIBILITY TO 7 AT AchzivR453Spring2020
AchzivR454Spring2020

---

### step 1 load data library etc and set wd


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(plyr)
library(ggplot2)
library(dplyr)
library(labdsv)
library(tidyverse)
library(reshape2)
library(vegan)
library(assertthat)
library(devtools)
#For Diversity calculations
library(plotrix)
library(betapart)
library(here)
library(ggridges)

library(ggtext)# for italics in the plot legend 
library(ggpubr) 
# Richness
library(rareNMtests)
library(mobr)
here<-here::here


```

```{r}
# Define the path to your CSV file using here()
csv_path <- here("data", "fishsurvey_sub2022.csv")
csv_path1 <- here("data", "monitoring_sites.csv")

# Read the CSV files

fish_data <- read.csv(csv_path)
sites <- read.csv(csv_path1)
complex.inic <-read.csv(here("data", "EnvVar10_45.csv"))

fish_data <-merge(fish_data,sites, by= "site_id") 

# Replace Sdot yam with SdotYam

fish_data$site[fish_data$site == 'Sdot yam'] <- 'SdotYam'
fish_data$season[fish_data$season == 'fall'] <- 'Fall'
fish_data = subset(fish_data, select = -c(tag) )
fish_data<- fish_data %>% mutate (tag= paste0(site,depth_category,transect,season,sample_year))
#Make sure Achziv is Reserve and Sdot Yam Not

head(fish_data)
```

###QA
```{r}

# Group by "tag" and count unique "sample_date" values
result <- fish_data %>%
  group_by(tag) %>%
  summarise(unique_dates = n_distinct(sample_date)) %>%
  filter(unique_dates > 1)

# Print the result
print(result)
```

### Subsetting DATA: to include data =="Achziv" \|site=="SdotYam sites, fish_filter will encompass data without fish empty transects and other outliers; all_data_clean includes fish empty transects and outliers and fill be used for richness and diversity calcs
Filtering cryptic species and species in the genus recognition level

```{r Subsetting DATA}
#selecting sublist for comparisons
fish_filter<-fish_data %>% filter(site=="Achziv"| site=="SdotYam" )
visibility_distribution <- table(fish_filter$visibility)
print(visibility_distribution)

fish_filter <-   fish_filter%>%
  filter(distance < 5.1)%>%
  filter(species != "Epinephelus sp") %>%
  filter(species != "Siganus sp") %>%
   filter(species != "Dasyatis.sp" ) %>%
  filter(species != "Empty transect") %>%
   filter(species != "Gobius sp" ) %>%
  filter(species != "Gobius bucchichi") %>%
  filter(species != "Gobius vittatus") %>%
   filter(species != "Dasyatis sp") 
fish_filter$reserve <- ifelse(fish_filter$site == "Achziv", "R", (ifelse(fish_filter$site_id == "SdotYam45", "N", "NR"))) # if its true write R if its false write Fall
head(fish_filter)

# Filter season
fish_filter <-   fish_filter#%>%
 # filter(season == "Fall")
```



##before we will convert the data to the wide format and start plotting we will add two more summary columns that will come to use later on. lets calculate the total number of individuals per survey and the total number of species.
#A. calculate the total number of individuals per species 
```{r}

fish_long <- fish_filter %>%
  dplyr::select(-transect,-a_b,-habitat, -depth, -factor_visibility, -observer, -size,-distance) %>%
  dplyr::relocate(species,.after = tag) %>%
   dplyr::mutate( occ =1 ) %>% 
 dplyr::group_by(across(site_id:species)) %>% 
 dplyr::summarise(total_ind = sum(occ))
```
#B. Calculate  the total number of species (n_species_sample)and total individuals per survey (n_ind_sample))

```{r}
fish_long<-fish_long %>%
  dplyr::group_by(tag)%>% 
  dplyr::mutate(n_species_sample = n_distinct(species),
          n_ind_sample= sum(total_ind)) %>%
  dplyr::ungroup()
head(fish_long)
```

###Finally, lets convert our data to wide format:


```{r}
fish_wide<- spread(fish_long,species,total_ind,fill = 0)
head(fish_wide)
colnames(fish_wide)


```

###making Fish abudance table filtering empty species

```{r}


fish_wide <- fish_wide[, colSums(fish_wide != 0) > 0]

write.csv(fish_wide, here("R.generated.Data", "45Fish_data_clean.csv"))

```

### Invertebrate summary according to taxonomy_group : allFish10.45dataFact


```{r}

Inv.Grp.data <- read.csv(here( "Data","inv_group_mat_wis.csv"))
Inv.Grp.data = subset(Inv.Grp.data, select = -c(tag) )
Inv.Grp.data<- Inv.Grp.data %>% mutate (tag= paste0(site
,depth,transect,season,year))
colnames(Inv.Grp.data)
Inv.Grp.factors <- dplyr::select(Inv.Grp.data,Algae,Bryozoan,Hard.Substrate, Hydrozoa, Soft.Substrate,Sponge,tag, 11:19)

head(Inv.Grp.factors)

DepthFishdataINv <- dplyr::select(Inv.Grp.factors, 1:7)


```
###Combine Rock cover wiskonsin transformed and complexity factors into the DepthFishdataFact df
```{r}

DepthFishdataFact <-left_join(fish_wide, DepthFishdataINv, by = "tag")
DepthFishdataFact <-left_join(DepthFishdataFact, complex.inic, by = "tag")

colnames(DepthFishdataFact)
DepthFishdataFact <-DepthFishdataFact %>%
  relocate("Algae" :"TempDiff",.after = tag) 
head(DepthFishdataFact)
colnames(DepthFishdataFact)


```
###Resolving date discrepencies between Fish and epibenthic cover data sets
```{r}



checktag<-dplyr::select(DepthFishdataFact,4,20,13)
#view(checktag)
# Update sample_date.x column where sample_date.y is not NA and is different
DepthFishdataFact$sample_date.x[!is.na(DepthFishdataFact$sample_date.y) & DepthFishdataFact$sample_date.y != DepthFishdataFact$sample_date.x] <- DepthFishdataFact$sample_date.y[!is.na(DepthFishdataFact$sample_date.y) & DepthFishdataFact$sample_date.y != DepthFishdataFact$sample_date.x]
#remove extra date column
DepthFishdataFact <- DepthFishdataFact[, !names(DepthFishdataFact) %in% "sample_date.y"]

# Print the updated dataframe to verify the changes
colnames(DepthFishdataFact)
DepthFishdataFact$sample_date <- DepthFishdataFact$sample_date.x
DepthFishdataFact$sample_date.x <- NULL
DepthFishdataFact <-DepthFishdataFact %>%
  relocate(sample_date,.after = sample_year) 
write.csv(DepthFishdataFact, here("R.generated.Data","Fish.FactorsGrp.wis.csv"))

```
####Diversity
```{r}

color6 = c("lightblue3", "blue2", "darkblue","yellow3", "darkorange", "brown")
my_data <- DepthFishdataFact %>%
  mutate(Year_season_id =paste0(sample_year,season,site_id)) %>%
  relocate("Year_season_id",.after = tag) 
 head(my_data)
 colnames(my_data)
```

###Species Matrix +Metadata +transformations
```{r}
colnames(my_data)
first_species <- 32
sp_matrix <- my_data[,first_species:ncol(my_data)] %>%
  select_if(~ sum(.) != 0)
meta_data <- my_data[,1:31] # columns i want to keep from my data


hel_trans <- decostand (sp_matrix, method = "hellinger" )
log10_trans_hel <- decostand (hel_trans,method = "log" ,base = 10)
log10_hel_my_data <- bind_cols(my_data[, -( first_species: ncol(my_data))], log10_trans_hel)
colnames (log10_hel_my_data)
```
### Hill numbers profile Allways use  hill = T
```{r}
renyi_profile <- renyi(sp_matrix,  
                       scales = c(0, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 4, 8, 16, 32, 64, Inf),
                       hill = T)
head(renyi_profile)

renyi_df<-bind_cols(meta_data,renyi_profile)

renyi_df<-gather(renyi_df,"Q","Value",30:44)
head(renyi_df)
renyi_df<-renyi_df %>% arrange(site_id,depth_category,sample_year, season,tag) %>%
   filter(!is.na(Q))

renyi_df$Q <- as.numeric(renyi_df$Q)
```

```{r}
head(renyi_df)

ggplot(data = renyi_df) +
  aes(x = Q, y = Value, group = tag, color = factor(sample_date)) +
  geom_line() +
  scale_x_continuous(limits = c(0, 10)) +
  scale_y_continuous(limits = c(0, 13)) +
  facet_wrap(. ~ site_id) +
  labs(y = expression("Diversity  D(p) "^"Q")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) +
 theme (legend.position = "none")



```

###SEction 2: Grouping per sampling day
```{r}
colnames(my_data)
group_year_data <- my_data %>% 
  dplyr::select(date_id, Year_season_id, site_id, depth_category, season, "Acantholabrus palloni" :"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni" :"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_year_data)
colnames(group_year_data)

colnames(log10_hel_my_data)
group_trans_data <- log10_hel_my_data %>% 
  dplyr::select(date_id, sample_year,site, Year_season_id, site_id, depth_category, season, "Acantholabrus palloni" :"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id, site, sample_year, site_id,depth_category, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni" :"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_trans_data)



```


```{r}

renyi_profile_group <- renyi(group_year_data[,-c(1:5)],
                             scales = c(0, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 4, 8, 16, 32, 64, Inf),  hill = T)

renyi_df_group<-bind_cols(group_year_data[,1:5],renyi_profile_group)

#compose long df with  new Q grouping
renyi_df_group<-gather(renyi_df_group,"Q","Value",6:20)

renyi_df_group<-renyi_df_group %>% arrange(Year_season_id, site_id,depth_category, season )

renyi_df_group$Q <- as.numeric(renyi_df_group$Q)
renyi_df_group$depth_category <- as.factor(renyi_df_group$depth_category)
head(renyi_df_group)
```
```{r}
ggplot(data = renyi_df_group) +
  aes(x = Q, y = Value, group = Year_season_id, color = factor(season)) +
  geom_line() +
  scale_x_continuous(limits = c(0, 10)) +
  scale_y_continuous(limits = c(0, 20)) +
  facet_wrap(. ~ site_id) +
  labs(y = expression("Diversity  D(p) "^"Q")) +
  theme_classic() +
    theme(legend.position = "none")
```


#



```{r}



bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(my_data$site_id)){
  
site_data <- my_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

NTTransectBeta <- ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6) # Specify color palette

NTTransectBeta
```

###Now with transect transformed data: log10_hel_my_data 
```{r}



bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(log10_hel_my_data$site_id)){
  
site_data <- log10_hel_my_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

TransectBeta <- ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6) # Specify color palette

TransectBeta
```
# grouped data

```{r}



bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(group_year_data$site_id)){
  
site_data <- group_year_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame



# lets plot:

pGr <-ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6) +theme_classic()  # Specify color palette +

pGr


```
###Now with transformed data: log10_hel_my_data grouped in group_trans_data 

```{r}
bray_scores_site<-list() # create an empty list to store my newly created data

for (i in unique(group_trans_data$site_id)){
  
site_data <- group_trans_data %>% filter(site_id == i) # keep only the observation of sample i

site_sp_matrix <- site_data[,first_species:length(site_data)] # create species matrix

site_bray_part <- bray.part(site_sp_matrix) # apply the function that calculate bray Curtis distances 
site_bray_results<- site_bray_part[[3]] # keep only the bray Curtis results

site_bray_results<-as.numeric(site_bray_results) # convert to numeric object

mean_bray <- mean(site_bray_results)
se_bray <- std.error(site_bray_results)
Site <- i

bray_data_site <- data.frame(Site,mean_bray,se_bray)

bray_scores_site[[i]]<-bray_data_site
  
}



bray_scores_site<-bind_rows(bray_scores_site) # convert from list to data frame
```
#grouped transformed

```{r}
PGrTrans <- ggplot(bray_scores_site,aes(x = Site,
                       y = mean_bray,
                       color = Site)) +
  geom_point(size = 4)+
  geom_errorbar(aes(ymin= mean_bray - se_bray,
                    ymax= mean_bray + se_bray),size =1.2,width = 0.2) +
  scale_color_manual(values = color6) + theme_classic()  +theme(axis.title.x = element_blank() ) +
           labs(y = "Grouped transformed Bray-Curtis dissimilarity")# Specify color palette

PGrTrans
```



```{r}
site_id_distribution <- table(group_year_data$site_id,group_year_data$season )
print(site_id_distribution)
```


```{r}
least_sampled <- 23

least_sampled
```



##Incidence data- beta pairs
```{r}
AC_incidence <- sp_matrix %>% 
  replace(sp_matrix > 0, 1) 
AC_incidence_bind<-bind_cols(meta_data,AC_incidence )


caves_beta_pairs <- beta.pair(AC_incidence)
```


##Group Incidence data- beta pairs
```{r}
head(group_year_data)
sp_data_group <- group_year_data %>%
  ungroup()  %>% 
  select(6:48)
meta_group<-group_year_data %>%
  ungroup()  %>% 
  select(1:5)
AC_incidenceGrp <- sp_data_group %>%
  replace(sp_data_group > 0, 1) 
AC_incidence<-bind_cols(meta_group,AC_incidenceGrp )


caves_beta_pairs <- beta.pair(AC_incidenceGrp)
```
```{r message=FALSE, results='hide'}


caves_beta_sample <- beta.sample(AC_incidenceGrp, 
            sites= 10,  # how many knolls to randomly sample
            samples = 10) # how many repetitions of (10) knolls to do 


```


after we understood how the function is working lets calculate the multi-site beta diversity of all the years for the site_id   
###Sorensen beta diversity: per transect 

```{r message=FALSE, warning=FALSE, results='hide'}


caves <-my_data  

caves[first_species:ncol(caves)] <-replace(caves[first_species:ncol(caves)],
                                            caves[first_species:ncol(caves)] > 0,
                                            1) # incident data

caves_all_years<-list() # empty list

for (i in unique(caves$site_id)) { # loop
  
  one_year <-  caves %>% filter(site_id == i)
  
  one_year_sp_matrix <- one_year[,first_species:ncol(one_year)]
  
  caves_beta_sample <- beta.sample(one_year_sp_matrix,
            sites= 5, 
            samples = 23)
  
 
  caves_beta_sample_data <- data.frame(mean = caves_beta_sample[["mean.values"]],
                                    se = caves_beta_sample[["sd.values"]])

  caves_beta_sample_data <-rownames_to_column(caves_beta_sample_data,'Component') 

  caves_beta_sample_data$Year_season <- i
  
 caves_all_years[[i]]<-caves_beta_sample_data
  
}


caves_all_years<-bind_rows(caves_all_years)


```


```{r}
caves_all_years<-caves_all_years %>% filter(Component != "beta.SOR")

ggplot(caves_all_years)+
  aes(x=Year_season,y=mean,fill=Component)+
  geom_bar(stat="identity",position = position_stack())+
  ylab(expression(paste(beta,"-diversity" (Sorensen)))) +theme_classic() +
  xlab("Site id")




```
###Sorensen beta diversity:  grouped

```{r message=FALSE, warning=FALSE, results='hide'}


caves <-group_year_data  
first_species = 6
caves[first_species:ncol(caves)] <-replace(caves[first_species:ncol(caves)],
                                            caves[first_species:ncol(caves)] > 0,
                                            1) # incident data

caves_all_years<-list() # empty list

for (i in unique(caves$site_id)) { # loop
  
  one_year <-  caves %>% filter(site_id == i)
  
  one_year_sp_matrix <- one_year[,first_species:ncol(one_year)]
  
  caves_beta_sample <- beta.sample(one_year_sp_matrix,
            sites= 5, 
            samples = 23)
  
 
  caves_beta_sample_data <- data.frame(mean = caves_beta_sample[["mean.values"]],
                                    se = caves_beta_sample[["sd.values"]])

  caves_beta_sample_data <-rownames_to_column(caves_beta_sample_data,'Component') 

  caves_beta_sample_data$Year_season <- i
  
 caves_all_years[[i]]<-caves_beta_sample_data
  
}


caves_all_years<-bind_rows(caves_all_years)


```


```{r}
caves_all_years<-caves_all_years %>% filter(Component != "beta.SOR")

ggplot(caves_all_years)+
  aes(x=Year_season,y=mean,fill=Component)+
  geom_bar(stat="identity",position = position_stack())+
  ylab(expression(paste(beta,"-diversity" (Sorensen)))) +theme_classic() +
  xlab("Site id")




```
We will use the function `rarefaction.sample` from the package `rareNMtests`. 

In this function the first argument is the data name,  the method argument `sample-size` means that the function will use incidence data and `q = 0` is used for species richness rarefaction.  

Now let's create sample based rarefaction curve for Fall 2018.  
#  Sample based  rarefraction and individual based 

```{r}
sample_based_rare<-list()

for (i in unique(my_data$site_id)) {

one_year<-my_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,first_species:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q = 0)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Richness","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

Richp <- ggplot(sample_based_rare,aes(x=Samples,y= Richness,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 

Richp
```






#Simpson
```{r}
sample_based_rare<-list()

for (i in unique(my_data$site_id)) {

one_year<-my_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,first_species:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q = 2)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","True.diversity","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

TrueDp <- ggplot(sample_based_rare,aes(x=Samples,y= True.diversity,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
TrueDp

```


```{r}
sample_based_rare<-list()

for (i in unique(my_data$site_id)) {

one_year<-my_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,first_species:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q = 1)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Effective.diversity","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

EffDp <- ggplot(sample_based_rare,aes(x=Samples,y= Effective.diversity,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
EffDp

```
##Grouped sample based rarefaction

```{r}
sample_based_rare<-list()

for (i in unique(group_year_data$site_id)) {

one_year<-group_year_data %>% filter(site_id == i)
year_sp_matrix <- one_year[,6:ncol(one_year)]

rarefaction <- rarefaction.sample(year_sp_matrix, method = "sample-size", q =0)

rarefaction$Year_month<-i

sample_based_rare[[i]]<- rarefaction
  
}

sample_based_rare<-bind_rows(sample_based_rare)

colnames(sample_based_rare)<-c("Samples","Richness","Site.id")
#colvec <- c("lightblue", "blue", "darkorange", "brown")   colors according to your group

EffDp <- ggplot(sample_based_rare,aes(x=Samples,y= Richness,color = Site.id))+geom_line(size = 1.2) + scale_color_manual(values = color6) +theme_classic() +theme(
    legend.title = element_blank(),  # Remove legend title
    legend.text = element_text(size = 12)  # Enlarge legend text
  ) 
EffDp

```
```{r warning=FALSE,results='hide'}
colnames(group_year_data)
Grpupedsp_matrix<-group_year_data %>%
  ungroup() %>% 
  filter(site_id== "Achziv10"| site_id=="Achziv25")%>%
  select(site_id,  "Acantholabrus palloni"  :ncol(group_year_data)) %>% arrange(site_id)

Grpupedsp_matrix<-as.data.frame(Grpupedsp_matrix) # make sure your data is defined as data.frame

eco_test <- EcoTest.sample(Grpupedsp_matrix[,-1], by= Grpupedsp_matrix$site_id,niter = 500, MARGIN=1,q=1)

```


```{r}
plot.new()
plot(eco_test)

```


```{r}
RarePlots <- ggarrange( Richp, EffDp,TrueDp,
          labels = c("Richness ", "Effective diveristy", "True diversity"),          ncol = 3, nrow = 1, common.legend = TRUE )  # Set a common x-axis title)
     
 RarePlots

```

###Group based individual rarefaction


group the data


```{r}


real_group<-my_data %>% dplyr::select(site_id,32:ncol(my_data))%>% 
 dplyr::group_by(site_id) %>% 
   dplyr::summarise(across(.fns = sum), .groups = "drop")
head(real_group)
```


###Group based individual rarefaction

```{r}


ind_based_rare<-list()


for (i in unique(real_group$site_id)) {

one_year<-real_group %>% filter(site_id == i)
year_sp_matrix <- one_year[,2:ncol(one_year)]

rarefaction <- rarefaction.individual(year_sp_matrix, method = "sample-size", q = 0)

rarefaction$site_id<-i

ind_based_rare[[i]]<- rarefaction
  
}

ind_based_rare<-bind_rows(ind_based_rare)

colnames(ind_based_rare)<-c("Individuals","Richness","site_id")

ggplot(ind_based_rare,aes(x=Individuals,y=Richness,color = site_id))+
  geom_line(size = 1.5)+ scale_color_manual(values = color6) + theme_classic()


```


```{r}
check<-real_group %>% mutate(total_abundance = rowSums(real_group[2:length(real_group)]),
         total_rihcness = rowSums(real_group[2:length(real_group)]>0)) %>% 
  select(site_id,total_abundance,total_rihcness)
  

```
###Ordinations on log base 10 due to siganiid densities

```{r}
library(ggfortify)
colnames(my_data)
#remove outlier Achziv 45 sample
Outlier_ac4516792 <- my_data %>%
  filter(date_id == "16792") %>%
   filter(depth_category >10)
my_data = my_data[!my_data$date_id == "16792",]
species_matrix <- my_data[,32:ncol(my_data)]
meta_data <- my_data[,1:31] # columns i want to keep from my data

log_trans_data <- decostand (species_matrix, method = "log",base = 10)
helge_trans_data <- decostand (species_matrix, method = "hellinger")
log10_trans_hel <- decostand (helge_trans_data, method = "log",base = 10)
```
#Group based ordinations
```{r}
colnames(group_trans_data)
first_species=8
group_trans_data = group_trans_data[!group_trans_data$date_id == "16792",]
species_matrix <- group_trans_data[,first_species:ncol(group_trans_data)]
meta_data <- group_trans_data[,1:7] # columns i want to keep from my data

log_trans_data <- decostand (species_matrix, method = "log",base = 10)
helge_trans_data <- decostand (species_matrix, method = "hellinger")
log10_trans_hel <- decostand (helge_trans_data, method = "log",base = 10)

```



###Dat preparation for RDA models A. complete epibentich +fish
```{r}
colnames(my_data)
group_year_data <- my_data %>% 
  dplyr::select(date_id, site, Year_season_id, site_id, depth_category,sample_year, season, "Acantholabrus palloni" :"Torquigener flavimaculosus") %>% 
  dplyr::group_by(date_id,site, site_id,depth_category,sample_year, season,Year_season_id) %>% # use this unit for analyses
  dplyr::summarise(across("Acantholabrus palloni" :"Torquigener flavimaculosus", sum),.groups = "keep")
head(group_year_data)
colnames(group_year_data)
rda_data <-  group_year_data

rda_data_complete   <- group_year_data[complete.cases(group_year_data), ]
rda_data_complete$site_id <-as.factor(rda_data_complete$site_id)
colnames(rda_data_complete)
species_rda_c <- rda_data_complete[,8:ncol(rda_data_complete)]
meta_rda_c <- rda_data_complete[,1:7] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_rda_c, method = "hellinger")
log10_trans_hel_c <- decostand (helge_trans_data_c, method = "log",base = 2)

colnames(rda_data_complete)
```
```{r}
rda_results <- rda(log10_trans_hel_c ~ site +season + depth_category+sample_year,
                data = rda_data_complete)

rda_results
summary (rda_results)
```
log10_trans_hel_c ~ season + depth_category+ reserve+ Hard.Substrate +Soft.Substrate + Algae ,
                data = rda_data_complete
                
#CCA was tested but it explains even less of the variation.
cca_results <- cca(log10_trans_hel~ season + depth_category+ reserve+ Hard.Substrate +Soft.Substrate + Algae ,
                data = my_data,
                na.action = na.omit) 

plot(cca_results, scaling = 3)

cca_results


```{r}
ordistep(rda_results ,na.action = na.omit)
```



Remember. When you see "Inertia", think "Variance". When you see "constrained" think "explained". So, in this example you can see that a proportion of 0.1121 of the inertia is explained. In words, the site_id,Site, knoll surface area and  depth affects ~10% of the all the community variance.


now lets plot the RDA using the `plot` function.
useful terminology for RDA plots:  

* sites = the scores (coordinates) or each row (i.e sample)    
* species = the unconstrained variables, actual species in our case    
* bp = bi plot, the effect of each constrained variable on the data    
* cn = centroid of non-continues constrained variable (for example Site or season)   
### RDA significance


The significance of your RDA can be tested using the function `anova.cca` which preform Anova like permutation tests. 

* don't be confused by the `cca` in the function name it can operate also on RDA

*Use either significance testing or model selection approch - not both!*
```{r}

ordiplot(rda_results, scaling = 2)
```

```{r}



colvec <- color6 #  colors according to your group



plot(rda_results,xlim = c(-3,3), 
     ylim = c(-3,3),
     # label the plot (title, and axes)
     main = "Biplot RDA ",
     xlab = paste0("RDA1 (16.2) %"), 
     ylab = paste0("RDA2 (5.2%)") ) # plot RDA

with(rda_data_complete, points(rda_results,
                     display = "sites", col = colvec[site_id],
                     pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores

with(rda_data_complete, legend("topleft", legend = levels(site_id), bty = "n",
                       col = colvec, pch = 21, pt.bg = colvec)) # add legend

#orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

text(rda_results,  display = "cn", cex = 1, col = "grey") #, add labels to centroieds


```

```{r}



colvec <- color6 #  colors according to your group



plot(rda_results,xlim = c(-3,3), 
     ylim = c(-3,3),
     # label the plot (title, and axes)
     main = "Biplot RDA ",
     xlab = paste0("RDA1 (16.2) %"), 
     ylab = paste0("RDA2 (5.2%)") ) # plot RDA

#with(rda_data_complete, points(rda_results,                      display = "sites", col = colvec[site_id],                     pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores

#with(rda_data_complete, legend("topleft", legend = levels(site_id), bty = "n",                       col = colvec, pch = 21, pt.bg = colvec)) # add legend

#orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

text(rda_results,  display = "cn", cex = 1, col = "Blue") #, add labels to centroieds


```


#The position of each species point in the ordination plot is determined by the strength and direction of the relationship between that species and the environmental variables included in the analysis. 
```{r}
plot(rda_results,xlim = c(-2,2), 
     ylim = c(-3,3),
     # label the plot (title, and axes)
     main = "Biplot RDA - scaling 2",
     xlab = paste0("RDA1 (10.8) %"), 
     ylab = paste0("RDA2 (4.2%)") ) # plot RDA

with(rda_data_complete, points(rda_results,
                     display = "sites", col = colvec[site_id],
                     pch = 21, bg = colvec[site_id]),cex=0.5) # add color to the scores

with(rda_data_complete, legend("bottomright", legend = levels(site_id), bty = "n",
                       col = colvec, pch = 21, pt.bg = colvec)) # add legend

#orditorp(rda_results,display = "species",choices = c(1, 2),air =1, col = "red") # add labels to some species

text(rda_results, display = "species", cex = 1, col = "blue") # add labels to centroieds


```
###Creating complete case data for temp diff and complexity measures
```{r}


my_data_complete   <- my_data[complete.cases(my_data), ]
colnames(my_data_complete)

my_data_complete <- my_data_complete[, colSums(my_data_complete != 0) > 0]
colnames(my_data_complete)
species_matrix_c <- my_data_complete[,first_species:ncol(my_data_complete)]
meta_data_c <- my_data_complete[,1:31] # columns i want to keep from my data
helge_trans_data_c <- decostand (species_matrix_c, method = "hellinger")
log10_trans_hel_c <- decostand (helge_trans_data_c, method = "log",base = 10)
```
```{r}

rda_results_CC <- rda(log10_trans_hel_c  ~  mean_depth + TempDiff + site +Rugosity +Curvature + season,
                data = meta_data_c)

rda_results_CC
```


```{r}
ordistep(rda_results_CC ,na.action = na.omit)
```
```{r}
summary (rda_results_CC)

```


###Pairwize tukey

```{r}
nmds_data<-vegdist(log10_trans_hel,method = "bray")

pairwise<-betadisper(nmds_data, # the distance matrix
                     group_year_data$site_id) # the groups
# Reset graphics device state before opening a new PNG device
graphics.off()

# Open a PNG device with a larger size
png("tukeyhsd_plot.png", width = 1000, height = 600, res = 100)

# Set margins and plot region
par(mar = c(5, 15, 4, 2) + 4)  # Increase left margin (second value)
par(plt = c(0.2, 0.95, 0.1, 0.9))  # Adjust plot region

# Plot the Tukey HSD results
plot(TukeyHSD(pairwise), las = 1, xpd = TRUE)

# Close the device
dev.off()







```


```{r}
ord  <- metaMDS(nmds_data,trace = FALSE)
```

  
We can see the Shepard plot of the nMDS. This described the "stress", Or, how well the nMDS is doing keeping the dissimilarity between points the same.  
 


*BTW - this plot is like assumption tests in models you should look at it but you don't need to present it in your final project...*

Large scatter around the line suggests that original dissimilarities are not well preserved in the reduced number of dimensions. You want to keep the stress as low as possible (anything above 0.2 is suboptimal).


Now let's check out the stress:

```{r}
ord$stress

```

OK, we preformed the analyses **now we can do 2 things:**

**1.** Visualize our results by plotting the nMDS  
**2.** Run statistic analyses

### 1. Visualize nMDS

Let's plot:

```{r }

plot(ord)

```

We can add some more information to the plot like different color for the different groups and convexhull 


* Note - you need ro run the lines together. Either mark all of them and than press `run` or `ctrl`+`enter` or (in r markdown) use the green arrow in the upper corner named -  `run current chunk`

```{r message=FALSE, warning=FALSE}

ordiplot(ord, type = "n",main = paste("stress=",round(ord$stress,3))) # create the plot
orditorp(ord, label = T, display = "sites" ,pch = 16) # add scores and color by sites
ordihull(ord, groups = my_data$site_id, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color6) # add convex hull
#legend("topleft", legend = levels(my_data$depth_category), bty = "n", col = colvec, pch = 15,cex=1.1) # add legend

```


the answer is Yes - Fish communities differ by site_id.  

if we want to know which survey were significantly different from one-another we can use post-hoc test and apply pair-wise comparison between the diffrent years.  

```{r}

pairwise<-betadisper(nmds_data, # the distance matrix
                     my_data$site_id) # the groups

par(mar=c(5,15,4,1))
plot(TukeyHSD(pairwise),las=2)

```

#### Simper

The simper functions performs pairwise comparisons of groups (i.e, `site_id`) and finds the contribution of each species to the average between-group Bray-Curtis dissimilarity.

**meaning** - which species contributed to the differences found between surveys.


```{r, results='hide'}
simper_output<-simper(log10_trans_hel, #my community data (transformed)
                      group = my_data$depth_category) # my group

summary(simper_output)
```

#Unique species by site
```{r}
site_id_fish <- fish_long %>% 
  filter(site =="Achziv" |site == "SdotYam")   
Td_fish<-  as.data.frame(table(site_id_fish$site_id, site_id_fish$species))%>%
  dplyr::filter(Freq>0)
 
 SdotYam45<- dplyr::filter(Td_fish,Var1 =="SdotYam45")

 SY45List <- as.character(unique(SdotYam45$Var2))
print(SY45List)

 SdotYam25<- dplyr::filter(Td_fish,Var1 =="SdotYam25")

 SY25List <- as.character(unique(SdotYam25$Var2))
print(SY25List)

```
###Trait data
```{r}
library(rfishbase)
library(dplyr)  # For data manipulation

# Your species list
SY45List <- c("Acantholabrus palloni", "Chromis chromis", "Coris julis", "Diplodus vulgaris", 
              "Epinephelus costae", "Epinephelus marginatus", "Muraena helena", "Mycteroperca rubra", 
              "Parupeneus forsskali", "Pterois miles", "Sparisoma cretense", "Sardin sp", 
              "Scomber sp", "Serranus cabrilla", "Siganus rivulatus", "Sparisoma cretense", 
              "Symphodus mediterraneus", "Symphodus tinca", "Synodus saurus", "Thalassoma pavo", 
              "Torquigener flavimaculosus")

# Function to safely get species info
get_species_info <- function(species_name) {
  tryCatch({
    info <- species(species_name)
    if (nrow(info) > 0) {
      return(info)
    } else {
      return(NULL)
    }
  }, error = function(e) {
    message("Error processing ", species_name, ": ", e$message)
    return(NULL)
  })
}

# Get info for SY45  species
species_info_list <- lapply(SY45List, get_species_info)

# Combine all data frames into one
SY45species_df <- do.call(rbind, species_info_list)
not.in.dist.df<- setdiff(SY45List,SY45species_df$Species) 
# Clean up the dataframe
SY45species_df <- SY45species_df %>%
  select(1, 28:33) %>%
  distinct()
SY45List <- c(
  "Atherinomorus forskalii", "Balistes capriscus",
  "Chromis chromis", "Coris julis",
  "Diplodus vulgaris", "Epinephelus costae",
  "Epinephelus marginatus", "Muraena helena",
  "Mycteroperca rubra", "Parablennius rouxi",
  "Parupeneus forsskali", "Pterois miles",
  "Sparisoma cretense", "Sardin sp",
  "Sargocentron rubrum", "Scomberomorus commerson",
  "Seriola dumerili", "Serranus cabrilla",
  "Serranus scriba", "Siganus luridus",
  "Siganus rivulatus", "Sparisoma cretense",
  "Symphodus mediterraneus", "Symphodus tinca",
  "Thalassoma pavo", "Torquigener flavimaculosus"
)
# Get info for SY25  species
species_info_list <- lapply(SY25List, get_species_info)

# Combine all data frames into one
SY25species_df <- do.call(rbind, species_info_list)
not.in.dist.df<- setdiff(SY25List,SY25species_df$Species) 
print(not.in.dist.df)
# Clean up the dataframe
SY25species_df <- SY25species_df %>%
  select(1, 28:33) %>%
  distinct()
# View the result
print(SY25species_df)

in.45.in.25<- setdiff(SY45List,SY25species_df$Species) 
print(in.45.in.25)
SYspecies_in.45not.in.25 <- lapply(in.45.in.25, get_species_info)
print(SYspecies_in.45not.in.25)
SIMPER_Ach45 <-c("Sciaena umbra","Diplodus sargus", "Scarus ghobban")
SIMPER_Ach45_info_list <- lapply(SIMPER_Ach45, get_species_info)
SIMPER_Ach45_info_list <- do.call(rbind, SIMPER_Ach45_info_list)
SIMPER_Ach45_info_list <- SIMPER_Ach45_info_list %>%
  select(1, 28:33) %>%
  distinct()
# View the result
print(SIMPER_Ach45_info_list)
```


```{r}
# Get info for SY45  species
species_info_list <- lapply(SY45List, get_species_info)

# Combine all data frames into one
SY45species_df <- do.call(rbind, species_info_list)
not.in.dist.df<- setdiff(SY45List,SY45species_df$Species) 
# Clean up the dataframe
SY45species_df <- SY45species_df %>%
  select(1, 28:33) %>%
  distinct()

# View the result
print(SY45species_df)


uniqueSY45 <-setdiff(SY45List,SY25List)
uniqueSY45
uniqueAch45 <-setdiff(listAch45,listAch25)


```

